### 2024-12-03 Manuscript_figures.Rmd
Added Manuscript_figures.Rmd and .html

---
title: "Figures_MSA-PD"
author: "Rasmus Rydbirk"
date: "03-12-2024"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Setup

```{r setup, message = F}
library(conos)
library(magrittr)
library(dplyr)
library(cacoa) # github.com/kharchenkolab/cacoa
library(sccore)
library(scHelper) # github.com/rrydbirk/scHelper
library(qs)
library(ggplot2)
library(ggsci)
library(cowplot)
library(ggpubr)
library(STRINGdb)
library(reshape2)
library(ggforce)
library(ggpmisc)
library(RColorBrewer)
library(CRMetrics)
library(ggrepel)
library(circlize)
library(ComplexHeatmap)
library(stringr)
library(rstatix)
library(slingshot)

## Comparisons
comp <- list(c("CTRL","MSA"),
             c("CTRL","PD"),
             c("MSA","PD"))

comp.long <- list(c("CTRL vs. MSA","CTRL vs. PD"),
                  c("CTRL vs. MSA","PD vs. MSA"),
                  c("CTRL vs. PD","PD vs. MSA"))

# Palettes

## Major celltypes
anno.neurons <- qread("anno_neurons.qs")
anno.neurons.major <- anno.neurons %>% 
  collapseAnnotation("GABA") %>% 
  collapseAnnotation("GLU") %>% 
  renameAnnotation("GABA", "Inh. neurons") %>% 
  renameAnnotation("GLU", "Exc. neurons") %>% 
  collapseAnnotation("MSN")

anno <- qread("anno_major.qs")

anno.major <- anno[!names(anno) %in% names(anno.neurons.major) & !anno == "Neurons"] %>% 
  {factor(c(., anno.neurons.major))}

# qsave(anno.major, "/work/msa/03_annotations/anno_0final.qs")

pal.major <- brewer.pal(n = 10, "Set1") %>% 
  c("lightblue3") %>% 
  setNames(levels(anno.major)[c(9,8,3,5,10,6:7,2,1,4)])

## Condition
pal.major %<>% c(c("yellow", brewer.pal(n = 6, "Accent")[5:6]) %>% setNames(c("CTRL","PD","MSA")))

## Comparisons
pal.major %<>% c(brewer.pal(n = 10, "Paired")[c(6,8,10)] %>% setNames(c("CTRL vs. MSA","CTRL vs. PD","PD vs. MSA")))

## Neurons
pal.major %<>% c(setNames(c("navy",
                            "mediumblue",
                            "lightslateblue",
                            "lightskyblue",
                            "lightblue",
                            "lightseagreen",
                            "blue",
                            "cadetblue1",
                            "cyan",
                            "cyan4",
                            "aquamarine2",
                            "lightcoral",
                            "brown1",
                            "orange",
                            "darkorange4",
                            "lightgoldenrod3"), 
                          levels(anno.neurons)))

## Glia
anno.glia <- c("anno_astro.qs",
               "anno_oligo.qs",
               "anno_opc.qs") %>% 
  lapply(qread) %>% 
  Reduce(c, .) %>% 
  factor() %>% 
  renameAnnotation("Homeostatic_astrocytes", "AS_homeostatic") %>% 
  renameAnnotation("Reactive_astrocytes", "AS_reactive") %>% 
  renameAnnotation("Homeostatic_LINC01608", "OL_LINC01608") %>%
  renameAnnotation("Homeostatic_SLC5A11", "OL_SLC5A11") %>% 
  renameAnnotation("Reactive_SCGZ", "OL_SGCZ")

pal.major %<>% c(setNames(c("pink",
                            pal.major["Astrocytes"],
                            pal.major["Oligodendrocytes"],
                            "chartreuse",
                            "darkolivegreen",
                            "brown4",
                            "coral"),
                          levels(anno.glia)))

## Microglia
anno.micro <- qread("anno_micro_pvm.qs") %>% 
  renameAnnotation("Steady-state","MIC_steady-state") %>% 
  renameAnnotation("Intermediate1","MIC_intermediate1") %>% 
  renameAnnotation("Intermediate2","MIC_intermediate2") %>% 
  renameAnnotation("Activated","MIC_activated")

pal.major %<>% c(setNames(c(pal.major["Microglia"],
                            "maroon4",
                            "magenta",
                            "pink3",
                            pal.major["PVMs"]),
                          levels(anno.micro)))

## Phago. assay
pal.major %<>% c(setNames(c("purple", 
                            "black",
                            pal.major[c("CTRL","PD","MSA")]), 
                          c("PBS",
                            "LPS",
                            "CTRL CSF",
                            "PD CSF",
                            "MSA CSF")))

## Houston assay
pal.major %<>% c(setNames(c(pal.major["PBS"],
                            pal.major["CTRL"],
                            "yellow3",
                            "orange",
                            pal.major["PD"],
                            "cyan4",
                            "navyblue",
                            pal.major["MSA"],
                            "pink3",
                            "red4"), 
                          c("Microglia medium",
                            "CTRL CSF, no dil.",
                            "CTRL CSF, 1:2 dil.",
                            "CTRL CSF, 1:4 dil.",
                            "PD CSF, no dil.",
                            "PD CSF, 1:2 dil.",
                            "PD CSF, 1:4 dil.",
                            "MSA CSF, no dil.",
                            "MSA CSF, 1:2 dil.",
                            "MSA CSF, 1:4 dil.")))

# Load major Conos object
con <- qread("con_major.qs", nthreads = 10)

tt <- Sys.time()
```

Define helper functions

```{r}
getOntWithFamily <- function(cao, comp) {
  fams <- cao$test.results[["GSEA"]]$families

  df.all <- list(cao$.__enclos_env__$private$getOntologyPvalueResults(name = "GSEA", genes = "up", p.adj = 1, q.value = 1),
                 cao$.__enclos_env__$private$getOntologyPvalueResults(name = "GSEA", genes = "down", p.adj = 1, q.value = 1)) %>% 
    bind_rows() %>% 
    mutate(logp = -log10(p.adjust),
           comp = comp)

  df <- cacoa:::getOntologyFamilyChildren(df.all, fams=fams, type = "GSEA", subtype = "BP")

  return(list(df.all = df.all,
              df = df,
              fams = fams))
}
```

```{r}
lianaCircos <- function(df,
                        top.interactions = 30, 
                        text.size = 1, 
                        pal = pal.major, 
                        cell.types = c("Astrocytes", 
                                       "Immune", 
                                       "Microglia",
                                       "Neurons",
                                       "OPCs",
                                       "Oligodendrocytes",
                                       "PVMs",
                                       "Pericytes/endothelial"),
                        big.gap = 5,
                        small.gap = 2,
                        arrow.width = 3,
                        link.ramp.rel = T,
                        link.sort = F,
                        scale = F,
                        arrow.head.width = 0.3,
                        arrow.head.length = 0.3,
                        link.ramp.col = c("navy", "grey", "firebrick")) {
  input_df <- df %>% 
    slice_max(order_by = score, n = top.interactions) %>% 
    mutate(target = paste0(target, " ")) %>% 
    mutate(source_lig = paste0(source, "|", ligand), 
           target_rec = paste0(target, "|", receptor))

  if (link.ramp.rel) {
    arr_wd <- rep(arrow.width, nrow(input_df))
  } else {
    arr_wd <- (((input_df$score - min(input_df$score))/(max(input_df$score) - min(input_df$score))) * (arrow.width)) + 1
  }

  # Colors and segments
  anno.col <- setNames(pal, 
                       cell.types) %>% 
    c(., "Oligodendrocytes " = unname(.["Oligodendrocytes"]))

  cell_cols <- anno.col[unique(c(unique(input_df$source), unique(input_df$target), "Oligodendrocytes "))]

  link_cols <- c()

  if (!link.ramp.rel) {
    for (i in input_df$source_lig) {
      link_cols <- c(link_cols, cell_cols[str_extract(i, 
                                                      "[^|]+")])
    }
  } else {
    input_df %<>% 
      arrange(score)

    df.down <- input_df %>% filter(score <= 0)
    link_down <- colorRampPalette(c(link.ramp.col[1], link.ramp.col[2]))(nrow(df.down))

    df.up <- input_df %>% filter(score > 0)
    link_up <- colorRampPalette(c(link.ramp.col[2], link.ramp.col[3]))(nrow(df.up))

    link_cols <- c(link_down, link_up)
  }

  segments <- unique(c(paste0(input_df$source, "|", input_df$ligand), 
                       paste0(input_df$target, "|", input_df$receptor)))

  grp <- str_extract(segments, "[^|]+") %>% 
    setNames(segments)

  # Redo colors
  cell_cols2 <- grp
  for (i in unique(grp)) {
    cell_cols2[cell_cols2 == i] <- cell_cols[i]
  }

  # Plot
  input_df %>% 
    select(source_lig, target_rec, score) %>%
    chordDiagram(directional = 1, 
                 group = grp,
                 scale = scale, 
                 diffHeight = 0.005, 
                 direction.type = c("arrows"),
                 link.arr.type = "triangle", 
                 annotationTrack = c(), 
                 preAllocateTracks = list(
                   list(track.height = 0.05),
                   list(track.height = 0.25),
                   list(track.height = 0.05)), 
                 big.gap = big.gap,
                 transparency = 1,
                 link.arr.lwd = arr_wd,
                 link.arr.col = link_cols,
                 link.arr.length = arrow.head.length,
                 link.arr.width = arrow.head.width,
                 small.gap = small.gap
    )

  circos.track(track.index = 2, panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, 
                CELL_META$ylim[1], 
                str_extract(CELL_META$sector.index, "[^|]+$"), 
                facing = "clockwise", 
                niceFacing = TRUE, 
                adj = c(0, 0.55), 
                cex = 1)
  }, bg.border = NA)

  # Split segments
  for (l in segments) {
    highlight.sector(l, track.index = 3, col = cell_cols2[l])
  }

  # Add ligand/receptor track
  ## Ligand
  highlight.sector(input_df$source_lig, 
                   track.index = 1, 
                   col = "black", 
                   text = "Ligands", 
                   cex = 1, 
                   text.col = "white", 
                   niceFacing = TRUE)
  ## Receptor
  highlight.sector(input_df$target_rec, 
                   track.index = 1, 
                   col = "white", 
                   text = "Receptors", 
                   cex = 1, 
                   text.col = "black", 
                   border = "black", 
                   niceFacing = TRUE)

  # Legends
  minmax <- input_df %>% 
    pull(score) %>% 
    {pmax(abs(min(.)), max(.))} %>% 
    formatC(digits = 1) %>% 
    as.numeric()

  col.range = c(-minmax, 0, minmax)
  lgd_links = Legend(at = col.range, 
                     col_fun = colorRamp2(col.range, link.ramp.col), 
                     title_position = "topleft", 
                     title = "Links")

  lgd_ct <- Legend(labels = unique(c(input_df$source, input_df$target)), 
                   title = "Cell type", 
                   type = "points",
                   legend_gp = gpar(col = "transparent"), 
                   background = cell_cols[unique(c(input_df$source, input_df$target))])

  lgd_list_vertical = packLegend(lgd_ct, lgd_links)

  draw(lgd_list_vertical, 
       just = c("left", "bottom"), 
       x = unit(5, "mm"), 
       y = unit(5, "mm"))

  circos.clear()
}
```

```{r}
getTscanTrajectory <- function(con, anno) {
  requireNamespace("TSCAN", quietly = T)

  emb <- con$embedding[names(anno), ]
  anno %<>% .[rownames(emb)]

  cent.ids <- emb %>% 
    rownames() %>% 
    split(anno)

  centroids <- cent.ids %>% 
    lapply(\(cid) emb[cid, ]) %>% 
    lapply(colMeans) %>% 
    bind_rows() %>% 
    t() %>% 
    `colnames<-`(c("UMAP1","UMAP2"))

  mst <- centroids %>% 
    TSCAN::createClusterMST(clusters = NULL)

  line.data <- TSCAN::reportEdges(centroids, mst = mst, clusters = NULL)

  return(line.data)
}
```

```{r, fig.width=8, fig.height=4}
plotGenePseudoBulk <- function(gene, cm.pseudo, legend = T) {
  idx <- cm.pseudo %>% 
    colnames() %>% 
    data.frame(id = .) %>% 
    mutate(condition = strsplit(id, "_|!!") %>% sget(1),
           ct = strsplit(id, "!!") %>% sget(2)) %>% 
    mutate(ord = order(condition, ct))

  x <- cm.pseudo %>% 
    .[match(gene, rownames(.)), match(colnames(na.omit(.)), colnames(.))] %>%
    .[idx$ord]

  plot.dat <- x %>% 
    {data.frame(sample = names(.), 
                value = unname(.))} %>%
    mutate(anno = strsplit(sample, "!!") %>% 
             sget(2),
           condition = strsplit(sample, "!!|_") %>% 
             sget(1)) %>%
    mutate(anno = factor(anno))

  stat.test <- plot.dat %>% 
    group_by(anno) %>% 
    rstatix::wilcox_test(value ~ condition) %>% 
    rstatix::add_xy_position(x = "anno", step.increase = 0.1)

  p <- plot.dat %>% 
    ggplot(aes(anno, value)) +
    geom_boxplot(aes(fill = condition)) +
    theme_bw() +
    theme(line = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          axis.ticks.x = element_line()) +
    labs(x = "", y = "Normalized pseudobulk expression", fill = "", title = paste0(gene, " expression")) +
    scale_fill_manual(values = ggsci::pal_jama()(3)) +
    stat_compare_means(aes(fill = condition), method = "kruskal", label = "p.signif") +
    stat_pvalue_manual(stat.test)

  if (!legend) p <- p + guides(fill = "none")

  return(p)
}
```

# Figure 1

## Figure 1b

```{r, fig.width=5.7, fig.height=4}
con$plotGraph(groups = anno.major, 
              embedding = "UMAP",
              size = 0.1, 
              palette = pal.major, 
              font.size = 3, 
              raster = T, 
              show.labels = T,
              plot.na = F, 
              show.legend = T,
              legend.title = "Cell type", 
              alpha = 0.05) + 
  dotSize(3) +
  labs(x="UMAP1", y= "UMAP2") +
  theme(line = element_blank())
```

## Figure 1c

```{r fig.height=24, fig.width=8}
c("RBFOX3","MOG","VCAN","AQP4","CSF1R","PDGFRB","PTPRC","MRC1","GAD1","SLC17A7","PPP1R1B") %>%
  sort() %>% 
  lapply(function(g) con$plotGraph(embedding = "UMAP", gene=g, title=g, size=0.1, plot.na = F, palette = colorRampPalette(c("grey90","firebrick"))) + theme(line = element_blank())) %>%
  cowplot::plot_grid(plotlist=., ncol=2)
```

## Figure 1d

```{r, fig.height=4, fig.width=5}
cm.merged <- con$getJointCountMatrix()

markers <- c("AQP4","PTPRC","CSF1R","RBFOX3","SLC17A7","GAD1","PPP1R1B","MOG","VCAN","PDGFRB","MRC1")

dotPlot(markers, 
        cm.merged, 
        anno.major %>% 
          factor(., levels = sort(levels(.))[c(1,3,5,2,4,6,7:10)]), 
        cols = c("white","firebrick"), 
        gene.order = markers)
```

# Figure 2

## Load data

```{r}
cao_msa <- qread("cao_major_msa.qs", nthreads = 10)
cao_pd <- qread("cao_major_pd.qs", nthreads = 10)
cao_dis <- qread("cao_major_dis.qs", nthreads = 10)
```

## Figure 2a

```{r}
spc <- con$getDatasetPerCell()

cpc <- spc %>% 
  as.character() %>% 
  grepl.replace(c("CTRL","MSA","PD")) %>% 
  `names<-`(names(spc))

con$plotGraph(groups = cpc, 
              embedding = "UMAP",
              size = 0.1, 
              alpha = 0.05,
              palette = pal.major, 
              font.size = 3, 
              raster = T, 
              show.labels = T,
              plot.na = F,
              mark.groups = F,
              show.legend = T) + 
  labs(x="UMAP1", y= "UMAP2", col = "") +
  theme(legend.position = "bottom",
        line = element_blank()) +
  dotSize(3)
```

## Figure 2b

```{r, fig.height=4, fig.width=3}
cao_msa$plotCellLoadings(show.pvals = F, 
                         alpha = 0)$data %>% 
  ggplot(aes(ind, values, fill = ind)) + 
  geom_hline(yintercept = 0, col = "black") +
  geom_violin() + 
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = pal.major) + 
  theme(legend.position = "none",
        line = element_blank()) +
  scale_y_continuous(breaks = c(-1,0,1), 
                     limits = c(-1,1),
                     labels = c("-1\nCTRL",0,"1\nMSA")) +
  labs(x = "", y = "")
```

## Figure 2c

```{r, fig.height=4, fig.width=3}
cao_pd$plotCellLoadings(show.pvals = F, 
                        alpha = 0)$data %>% 
  ggplot(aes(ind, values, fill = ind)) + 
  geom_hline(yintercept = 0, col = "black") +
  geom_violin() + 
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = pal.major) + 
  theme(legend.position = "none",
        line = element_blank()) +
  scale_y_continuous(breaks = c(-1,0,1), 
                     limits = c(-1,1),
                     labels = c("-1\nCTRL",0,"1\nPD")) +
  geom_vline(xintercept = 7.5, col = "red") +
  labs(x = "", y = "")
```

## Figure 2d

```{r, fig.height=4, fig.width=3}
cao_dis$plotCellLoadings(show.pvals = F, 
                         alpha = 0)$data %>% 
  ggplot(aes(ind, values, fill = ind)) + 
  geom_hline(yintercept = 0, col = "black") +
  geom_violin() + 
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = pal.major) + 
  theme(legend.position = "none",
        line = element_blank()) +
  scale_y_continuous(breaks = c(-1,0,1), 
                     limits = c(-1,1),
                     labels = c("-1\nPD",0,"1\nMSA")) +
  geom_vline(xintercept = 5.5, col = "red") +
  labs(x = "", y = "")
```

## Figure 2e

```{r, fig.width = 6, fig.height = 4}
cao_msa$cell.groups.palette <- pal.major
cao_pd$cell.groups.palette <- pal.major
cao_dis$cell.groups.palette <- pal.major

cao_msa$plotExpressionShiftMagnitudes() + labs(y = "Normalized expression distance")
cao_pd$plotExpressionShiftMagnitudes() + labs(y = "Normalized expression distance")
cao_dis$plotExpressionShiftMagnitudes() + labs(y = "Normalized expression distance")
```

# Figure 3

## Load data

```{r}
cao <- qread("cao_neurons.qs", nthreads = 10)
cao_msa <- qread("cao_neurons_msa.qs", nthreads = 10)
cao_pd <- qread("cao_neurons_pd.qs", nthreads = 10)
cao_dis <- qread("cao_neurons_dis.qs", nthreads = 10)
```

## Figure 3a

```{r, fig.height=4, fig.width=4}
cao$plotEmbedding(groups = cao$cell.groups,
                  size = 0.1, 
                  palette = pal.major, 
                  font.size = 3, 
                  raster = T, 
                  show.labels = T,
                  plot.na = F,
                  alpha = 0.1) + 
  theme(line = element_blank()) +
  labs(x="largeVis1", y= "largeVis2")
```

## Figure 3b

```{r, fig.width=6.7, fig.height=4}
cm.merged <- cao$data.object$getJointCountMatrix()

markers <- c("GAD1","GAD2","PPP1R1B","SLC17A7","CHAT","CALB2","DCC","ID2","MEIS2","ST18","NKX2-1","NR2F2","PAX6","PVALB","RXFP1","SST","VIP","RELN","SATB2","DRD1","DRD2","FOXP2", "LRP8", "VLDLR")

dotPlot(markers, 
        cm.merged, 
        anno.neurons, 
        cols = c("white","firebrick"), 
        gene.order = markers)
```

## Figure 3c

```{r, fig.height=4, fig.width=3}
cao_msa$plotCellLoadings(show.pvals = F, 
                         alpha = 0)$data %>% 
  ggplot(aes(ind, values, fill = ind)) + 
  geom_hline(yintercept = 0, col = "black") +
  geom_violin() + 
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = pal.major) + 
  theme(legend.position = "none",
        line = element_blank()) +
  scale_y_continuous(breaks = c(-1,0,1), 
                     limits = c(-1,1),
                     labels = c("-1\nCTRL",0,"1\nMSA")) +
  labs(x = "", y = "")
```

## Figure 3d

```{r, fig.height=4, fig.width=3}
cao_pd$plotCellLoadings(show.pvals = F, 
                        alpha = 0)$data %>% 
  ggplot(aes(ind, values, fill = ind)) + 
  geom_hline(yintercept = 0, col = "black") +
  geom_violin() + 
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = pal.major) + 
  theme(legend.position = "none",
        line = element_blank()) +
  scale_y_continuous(breaks = c(-1,0,1), 
                     limits = c(-1,1),
                     labels = c("-1\nCTRL",0,"1\nPD")) +
  geom_vline(xintercept = 13.5, col = "red") +
  labs(x = "", y = "")
```

## Figure 3e

```{r, fig.height=4, fig.width=3}
cao_dis$plotCellLoadings(show.pvals = F, 
                         alpha = 0)$data %>% 
  ggplot(aes(ind, values, fill = ind)) + 
  geom_hline(yintercept = 0, col = "black") +
  geom_violin() + 
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = pal.major) + 
  theme(legend.position = "none",
        line = element_blank()) +
  scale_y_continuous(breaks = c(-1,0,1), 
                     limits = c(-1,1),
                     labels = c("-1\nPD",0,"1\nMSA")) +
  labs(x = "", y = "")
```

## Figure 3f

```{r, fig.height=4, fig.width=4}
sample.groups <- cao$sample.groups

cao$plotCellDensity(show.cell.groups = F)$CTRL +
  geom_circle(aes(x0 = 0.75, y0 = -4.5, r = 1)) +
  geom_circle(aes(x0 = 3.8, y0 = -1.1, r = 0.5), col = "cyan3") +
  theme(line = element_blank())

# MSA
cao$sample.groups <- sample.groups %>% 
  names() %>% 
  grepl.replace(c("CTRL","MSA","PD")) %>% 
  `names<-`(sample.groups %>% names()) %>% 
  .[. %in% c("CTRL","MSA")]
cao$target.level <- "MSA"
cao$estimateCellDensity(method = "graph", name = "msa.density")
cao$estimateDiffCellDensity(type = "subtract", name = "msa.density")
cao$plotCellDensity(show.cell.groups = F, name = "msa.density")$MSA +
  geom_circle(aes(x0 = 0.75, y0 = -4.5, r = 1)) +
  geom_circle(aes(x0 = 3.8, y0 = -1.1, r = 0.5), col = "cyan3") +
  theme(line = element_blank())

# PD
cao$sample.groups <- sample.groups %>%
  names() %>%
  grepl.replace(c("CTRL","MSA","PD")) %>%
  `names<-`(sample.groups %>% names()) %>%
  .[. %in% c("CTRL","PD")]
cao$target.level <- "PD"
cao$estimateCellDensity(method = "graph", name = "pd.density")
cao$estimateDiffCellDensity(type = "subtract", name = "pd.density")
cao$plotCellDensity(show.cell.groups = F, name = "pd.density")$PD +
  geom_circle(aes(x0 = 0.75, y0 = -4.5, r = 1)) +
  geom_circle(aes(x0 = 3.8, y0 = -1.1, r = 0.5), col = "cyan3") +
  theme(line = element_blank())
```

## Figure 3g

```{r, fig.width = 8.5, fig.height = 5}
cao_msa$plotExpressionShiftMagnitudes() + labs(y = "Normalized expression distance")
cao_pd$plotExpressionShiftMagnitudes() + labs(y = "Normalized expression distance")
cao_dis$plotExpressionShiftMagnitudes() + labs(y = "Normalized expression distance")
```

# Figure 4

## Load data

```{r}
con.glia <- qread("/work/msa/02_data/conos/con_oligo_astro_opc.qs", nthreads = 10)

cao_msa <- qread("cao_oligo_astro_opc_msa.qs", nthreads = 10)
cao_pd <- qread("cao_oligo_astro_opc_pd.qs", nthreads = 10)
cao_dis <- qread("cao_oligo_astro_opc_dis.qs", nthreads = 10)
```

## Figure 4a

```{r, fig.width=4, fig.height=4}
con.glia$plotGraph(groups = anno.glia, 
                   plot.na = F, 
                   size = 0.1,
                   palette = pal.major, 
                   font.size = 3, 
                   raster = T, 
                   show.labels = T, embedding = "largeVis_CPCA_AS01",
                   alpha = 0.05) + 
  labs(x = "largeVis1", y = "largeVis2") +
  theme(line = element_blank())
```

## Figure 4b

```{r, fig.height=4, fig.width=4.5}
cm.merged <- con.glia$getJointCountMatrix()

markers <- c("AQP4","TNC","MOG","LINC01608","SLC5A11","SGCZ","VCAN","OLIG2")

dotPlot(markers, 
        cm.merged, 
        anno.glia,
        cols = c("white","firebrick"), 
        gene.order = markers)
```

## Figure 4c

```{r, fig.height=4, fig.width=3}
cao_msa$plotCellLoadings(show.pvals = F, 
                         alpha = 0)$data %>% 
  ggplot(aes(ind, values, fill = ind)) +
  geom_hline(yintercept = 0, col = "black") +
  geom_violin() + 
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = pal.major) + 
  theme(legend.position = "none",
        line = element_blank()) +
  scale_y_continuous(breaks = c(-1,0,1), 
                     limits = c(-1,1),
                     labels = c("-1\nCTRL",0,"1\nMSA")) +
  geom_vline(xintercept = 6.5, col = "red") +
  labs(x = "", y = "")
```

## Figure 4d

```{r, fig.height=4, fig.width=3}
cao_pd$plotCellLoadings(show.pvals = F, 
                        alpha = 0)$data %>% 
  ggplot(aes(ind, values, fill = ind)) + 
  geom_hline(yintercept = 0, col = "black") +
  geom_violin() + 
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = pal.major) + 
  theme(legend.position = "none",
        line = element_blank()) +
  scale_y_continuous(breaks = c(-1,0,1), 
                     limits = c(-1,1),
                     labels = c("-1\nCTRL",0,"1\nPD")) +
  geom_vline(xintercept = 4.5, col = "red") +
  labs(x = "", y = "")
```

## Figure 4e

```{r, fig.height=4, fig.width=3}
cao_dis$plotCellLoadings(show.pvals = F, 
                         alpha = 0)$data %>% 
  ggplot(aes(ind, values, fill = ind)) + 
  geom_hline(yintercept = 0, col = "black") +
  geom_violin() + 
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = pal.major) + 
  theme(legend.position = "none",
        line = element_blank()) +
  scale_y_continuous(breaks = c(-1,0,1), 
                     limits = c(-1,1),
                     labels = c("-1\nPD",0,"1\nMSA")) +
  geom_vline(xintercept = 6.5, col = "red") +
  labs(x = "", y = "")
```

## Figure 4f

```{r, fig.width = 7, fig.height = 5}
cao_msa$plotExpressionShiftMagnitudes() + labs(y = "Normalized expression distance")
cao_pd$plotExpressionShiftMagnitudes() + labs(y = "Normalized expression distance")
cao_dis$plotExpressionShiftMagnitudes() + labs(y = "Normalized expression distance")
```

## Figure 4h,i

```{r, fig.width=14, fig.height=5}
cao_msa <- qread("cao_astro_msa.qs", nthreads = 10)
cao_pd <- qread("cao_astro_pd.qs", nthreads = 10)
cao_dis <- qread("cao_astro_dis.qs", nthreads = 10)

df.all <- Map(getOntWithFamily, list(cao_msa, cao_pd, cao_dis), c("CTRL vs. MSA", "CTRL vs. PD", "PD vs. MSA")) %>% 
  lget("df.all") %>% 
  bind_rows() %>% 
  mutate(Group = paste0("AS_", strsplit(Group, "_") %>% sget(1) %>% tolower()))

# Select relevant pathways
p.ids <- c("GO:0050821", "GO:0050728", "GO:0006986", "GO:0030168", "GO:0030001", "GO:1904707", "GO:0140053", "GO:0006486", "GO:0010717", "GO:0031113", "GO:0032675", "GO:0061041", "GO:0071222", "GO:0045766", "GO:0006487", "GO:0097501", "GO:0042776", "GO:0045047", "GO:0044794", "GO:0035633", "GO:0097242", "GO:0006120", "GO:0042026", "GO:0006979", "GO:0071222", "GO:0017015", "GO:0009611", "GO:1901201", "GO:0042594", "GO:0097501", "GO:0006123", "GO:0006909", "GO:0007179", "GO:0006120", "GO:0097501")
```

### Figure 4h

```{r, fig.width=14, fig.height=3}
ct <- "AS_homeostatic"

df.sel <- df.all %>% 
  filter(Group == ct)

p1 <- ggplot(df.sel, aes(NES, logp, col = comp, shape = comp)) +
  geom_point(data = ~filter(., p.adjust > 0.05), size = 0.1, alpha = 0.5, col = "black") +
  geom_point(data = ~filter(., p.adjust <= 0.05), size = 2, alpha = 0.8) +
  theme_bw() +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  scale_color_manual(values = pal.major) +
  theme(line = element_blank()) +
  labs(y = "-log10(adj. p)", shape = "Comparison", col = "Comparison") +
  dotSize(3)

df <- Map(getOntWithFamily, list(cao_msa, cao_pd, cao_dis), c("CTRL vs. MSA", "CTRL vs. PD", "PD vs. MSA")) %>% 
  lget("df") %>% 
  bind_rows() %>% 
  mutate(Group = paste0("AS_", strsplit(Group, "_") %>% sget(1) %>% tolower())) %>% 
  filter(Group == ct,
         ID %in% p.ids)

p2 <- df %>% 
  filter(p.adjust <= 0.05, NES > 0) %>% 
  mutate(Group = factor(Group, levels = sort(unique(Group)))) %>% 
  select(Description, NES, Group, comp) %>%
  as.data.frame() %>% 
  group_by(comp, Group) %>% 
  group_split() %>% 
  lapply(dplyr::slice, seq(5)) %>%
  bind_rows() %>% 
  mutate(., x = 0, y = rev(seq(nrow(.)))) %>% 
  ggplot(aes(x, y, label = Description, col = comp)) + 
  geom_text(hjust = 0) +
  theme_void() +
  xlim(0, 1) +
  labs(title = "Selected terms for up-regulated genes") +
  scale_color_manual(values = pal.major) +
  guides(col = "none")

p3 <- df %>% 
  filter(p.adjust <= 0.05, NES < 0) %>% 
  select(Description, NES, Group, comp) %>%
  group_by(comp, Group) %>% 
  group_split() %>% 
  lapply(dplyr::slice, seq(5)) %>% 
  bind_rows() %>% 
  mutate(., x = 0, y = rev(seq(nrow(.)))) %>% 
  ggplot(aes(x, y, label = Description, col = comp)) + 
  geom_text(hjust = 0) +
  theme_void() +
  xlim(0, 1) +
  labs(title = "Selected terms for down-regulated genes") +
  scale_color_manual(values = pal.major) +
  guides(col = "none")

plot_grid(plotlist = list(p3, p1, p2), ncol = 3, rel_widths = c(0.9, 1, 0.9))
```

### Figure 4i

```{r, fig.width=14, fig.height=3}
ct <- "AS_reactive"

df.sel <- df.all %>% 
  filter(Group == ct)

p1 <- ggplot(df.sel, aes(NES, logp, col = comp, shape = comp)) +
  geom_point(data = ~filter(., p.adjust > 0.05), size = 0.1, alpha = 0.5, col = "black") +
  geom_point(data = ~filter(., p.adjust <= 0.05), size = 2, alpha = 0.8) +
  theme_bw() +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  scale_color_manual(values = pal.major) +
  theme(line = element_blank()) +
  labs(y = "-log10(adj. p)", shape = "Comparison", col = "Comparison") +
  dotSize(3)

df <- Map(getOntWithFamily, list(cao_msa, cao_pd, cao_dis), c("CTRL vs. MSA", "CTRL vs. PD", "PD vs. MSA")) %>% 
  lget("df") %>% 
  bind_rows() %>% 
  mutate(Group = paste0("AS_", strsplit(Group, "_") %>% sget(1) %>% tolower())) %>% 
  filter(Group == ct,
         ID %in% p.ids)

p2 <- df %>% 
  filter(p.adjust <= 0.05, NES > 0) %>% 
  mutate(Group = factor(Group, levels = sort(unique(Group)))) %>% 
  select(Description, NES, Group, comp) %>%
  group_by(comp, Group) %>% 
  group_split() %>% 
  lapply(dplyr::slice, seq(5)) %>%
  bind_rows() %>% 
  mutate(., x = 0, y = rev(seq(nrow(.)))) %>% 
  ggplot(aes(x, y, label = Description, col = comp)) + 
  geom_text(hjust = 0) +
  theme_void() +
  xlim(0, 1) +
  labs(title = "Selected terms for up-regulated genes") +
  scale_color_manual(values = pal.major) +
  guides(col = "none")

p3 <- df %>% 
  filter(p.adjust <= 0.05, NES < 0) %>% 
  select(Description, NES, Group, comp) %>%
  group_by(comp, Group) %>% 
  group_split() %>% 
  lapply(dplyr::slice, seq(5)) %>% 
  bind_rows() %>% 
  mutate(., x = 0, y = rev(seq(nrow(.)))) %>% 
  ggplot(aes(x, y, label = Description, col = comp)) + 
  geom_text(hjust = 0) +
  theme_void() +
  xlim(0, 1) +
  labs(title = "Selected terms for down-regulated genes") +
  scale_color_manual(values = pal.major) +
  guides(col = "none")

plot_grid(plotlist = list(p3, p1, p2), ncol = 3, rel_widths = c(0.9, 1, 0.9))
```

## Figure 4j

```{r, fig.width=14, fig.height=3}
cao_msa <- qread("cao_oligo_msa.qs", nthreads = 10)
cao_pd <- qread("cao_oligo_pd.qs", nthreads = 10)
cao_dis <- qread("cao_oligo_dis.qs", nthreads = 10)

df.all <- Map(getOntWithFamily, list(cao_msa, cao_pd, cao_dis), c("CTRL vs. MSA", "CTRL vs. PD", "PD vs. MSA")) %>% 
  lget("df.all") %>% 
  bind_rows() %>% 
  mutate(Group = paste0("OL_", strsplit(Group, "_") %>% sget(2) %>% gsub("SCGZ", "SGCZ", .)))

# Select relevant pathways
p.ids <- c("GO:0071345", "GO:0045596", "GO:0060271", "GO:0048709", "GO:0006986", "GO:0016032", "GO:0060271", "GO:0042026", "GO:0006986", "GO:0120034", "GO:0071346", "GO:1904262", "GO:0042776", "GO:0061077", "GO:0050821", "GO:0045047", "GO:0042776", "GO:0050821", "GO:0045047", "GO:0007042", "GO:0042776", "GO:0061077", "GO:0045047", "GO:0042776", "GO:0045047", "GO:2000249", "GO:0002474", "GO:0042026", "GO:0002040", "GO:0046330", "GO:0043542", "GO:0030036", "GO:0032981", "GO:0007042")

ct <- "OL_SLC5A11"

df.sel <- df.all %>% 
  filter(Group == ct)

p1 <- ggplot(df.sel, aes(NES, logp, col = comp, shape = comp)) +
  geom_point(data = ~filter(., p.adjust > 0.05), size = 0.1, alpha = 0.5, col = "black") +
  geom_point(data = ~filter(., p.adjust <= 0.05), size = 2, alpha = 0.8) +
  theme_bw() +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  scale_color_manual(values = pal.major) +
  theme(line = element_blank()) +
  labs(y = "-log10(adj. p)", shape = "Comparison", col = "Comparison") +
  dotSize(3)

df <- Map(getOntWithFamily, list(cao_msa, cao_pd, cao_dis), c("CTRL vs. MSA", "CTRL vs. PD", "PD vs. MSA")) %>% 
  lget("df") %>% 
  bind_rows() %>% 
  mutate(Group = paste0("OL_", strsplit(Group, "_") %>% sget(2) %>% gsub("SCGZ", "SGCZ", .))) %>% 
  filter(Group == ct,
         ID %in% p.ids)

p2 <- df %>% 
  filter(p.adjust <= 0.05, NES > 0) %>% 
  mutate(Group = factor(Group, levels = sort(unique(Group)))) %>% 
  select(Description, NES, Group, comp) %>%
  group_by(comp, Group) %>% 
  group_split() %>% 
  lapply(dplyr::slice, seq(5)) %>%
  bind_rows() %>% 
  mutate(., x = 0, y = rev(seq(nrow(.)))) %>% 
  ggplot(aes(x, y, label = Description, col = comp)) + 
  geom_text(hjust = 0) +
  theme_void() +
  xlim(0, 1) +
  labs(title = "Selected terms for up-regulated genes") +
  scale_color_manual(values = pal.major) +
  guides(col = "none")

p3 <- df %>% 
  filter(p.adjust <= 0.05, NES < 0) %>% 
  select(Description, NES, Group, comp) %>%
  group_by(comp, Group) %>% 
  group_split() %>% 
  lapply(dplyr::slice, seq(5)) %>% 
  bind_rows() %>% 
  mutate(., x = 0, y = rev(seq(nrow(.)))) %>% 
  ggplot(aes(x, y, label = Description, col = comp)) + 
  geom_text(hjust = 0) +
  theme_void() +
  xlim(0, 1) +
  labs(title = "Selected terms for down-regulated genes") +
  scale_color_manual(values = pal.major) +
  guides(col = "none")

plot_grid(plotlist = list(p3, p1, p2), ncol = 3, rel_widths = c(0.9, 1, 0.9))
```

## Figure 4k,l

```{r}
cm.merged <- con$getJointCountMatrix(raw = T) %>% 
  Matrix::t() %>% 
  .[, colnames(.) %in% names(anno.glia)]

# Create sample-wise annotation
anno.donor <- con$getDatasetPerCell()[colnames(cm.merged)]
anno.subtype <- anno.glia %>% 
  .[!is.na(.)] %>% 
  factor()

idx <- intersect(anno.donor %>% names(), anno.subtype %>% names())

anno.donor %<>% .[idx]
anno.subtype %<>% .[idx] %>% 
  {`names<-`(as.character(.), names(.))}

anno.final <- paste0(anno.donor,"!!",anno.subtype) %>% 
  `names<-`(anno.donor %>% names())

# Create pseudo CM
cm.pseudo <- sccore::collapseCellsByType(cm.merged %>% Matrix::t(), 
                                         groups = anno.final, min.cell.count = 1) %>% 
  t() %>%
  apply(2, as, "integer")

cm.pseudo %<>% 
  {. + 1} %>% 
  DESeq2::DESeqDataSetFromMatrix(., 
                                 colnames(.) %>% 
                                   strsplit("_") %>% 
                                   sget(1) %>% 
                                   data.frame() %>% 
                                   `dimnames<-`(list(colnames(cm.pseudo), "group")), 
                                 design = ~ group) %>% 
  DESeq2::estimateSizeFactors() %>% 
  DESeq2::counts(normalized = T)
```

### Figure 4k

```{r, fig.width=4, fig.height=4}
# For DE between visits
genes <- "TLR1"

idx <- cm.pseudo %>% 
  colnames() %>% 
  data.frame(id = .) %>% 
  mutate(condition = strsplit(id, "_|!!") %>% sget(1),
         ct = strsplit(id, "!!") %>% sget(2)) %>% 
  mutate(ord = order(condition, ct))

x <- cm.pseudo %>% 
  .[match(genes, rownames(.)), match(colnames(na.omit(.)), colnames(.))] %>%
  .[idx$ord]

plot.dat <- x %>% 
  {data.frame(sample = names(.), 
              value = unname(.))} %>%
  mutate(anno = strsplit(sample, "!!") %>% 
           sget(2),
         condition = strsplit(sample, "!!|_") %>% 
           sget(1)) %>% 
  filter(anno %in% c("OL_LINC01608", "OL_SGCZ", "OL_SLC5A11")) %>% 
  mutate(anno = factor(anno))

stat.test <- plot.dat %>% 
  group_by(anno) %>% 
  rstatix::wilcox_test(value ~ condition) %>% 
  rstatix::add_xy_position(x = "anno", step.increase = 0.1, fun = "mean_sd", scales = "free")

plot.dat %>% 
  ggplot(aes(anno, value)) +
  geom_boxplot(aes(fill = condition)) +
  theme_bw() +
  theme(line = element_blank(),
        axis.ticks.x = element_line(),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  labs(x = "", y = "Normalized pseudobulk expression", fill = "", title = "TLR1 expression") +
  scale_fill_manual(values = ggsci::pal_jama()(3)) +
  stat_compare_means(aes(fill = condition), method = "kruskal", label = "p.signif", label.y = 8.5e2) +
  stat_pvalue_manual(stat.test)
```

### Figure 4l

```{r, fig.width=4, fig.height=4}
# For DE between visits
genes <- "LRP1"

idx <- cm.pseudo %>% 
  colnames() %>% 
  data.frame(id = .) %>% 
  mutate(condition = strsplit(id, "_|!!") %>% sget(1),
         ct = strsplit(id, "!!") %>% sget(2)) %>% 
  mutate(ord = order(condition, ct))

x <- cm.pseudo %>% 
  .[match(genes, rownames(.)), match(colnames(na.omit(.)), colnames(.))] %>%
  .[idx$ord]

plot.dat <- x %>% 
  {data.frame(sample = names(.), 
              value = unname(.))} %>%
  mutate(anno = strsplit(sample, "!!") %>% 
           sget(2),
         condition = strsplit(sample, "!!|_") %>% 
           sget(1)) %>% 
  filter(anno %in% c("OL_LINC01608", "OL_SGCZ", "OL_SLC5A11")) %>% 
  mutate(anno = factor(anno))

stat.test <- plot.dat %>% 
  group_by(anno) %>% 
  rstatix::wilcox_test(value ~ condition) %>% 
  rstatix::add_xy_position(x = "anno", step.increase = 0.1)

plot.dat %>% 
  ggplot(aes(anno, value)) +
  geom_boxplot(aes(fill = condition)) +
  theme_bw() +
  theme(line = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.ticks.x = element_line()) +
  labs(x = "", y = "Normalized pseudobulk expression", fill = "", title = "LRP1 expression") +
  scale_fill_manual(values = ggsci::pal_jama()(3)) +
  stat_compare_means(aes(fill = condition), method = "kruskal", label = "p.signif", label.y = 175) +
  stat_pvalue_manual(stat.test) +
  ylim(c(0, 180))
```

## Figure 4m

For calculation of data, see `Liana.ipynb`.

```{r, fig.width=10, fig.height=8}
dat <- read.delim("liana_res.csv",
                  sep = ",",
                  header = T) %>%
  mutate(group = strsplit(sample, "_") %>%
           sget(1))

dat.plot <- dat %>% 
  dplyr::rename(ligand = ligand_complex,
         receptor = receptor_complex,
         score = lrscore) %>%
  filter(target == "Oligodendrocytes",
         ligand %in% c("BMP1", "SERPINE2", "PSAP", "APOE", "APP", "SERPING1"),
         receptor %in% c("LRP1", "BMPR1A", "LRP4")) %>%
  group_by(group, ligand, receptor, source, target) %>% 
  summarize(score = mean(score)) %>% 
  ungroup() %>% 
  arrange(group, source, target, ligand, receptor)

dat.msa <- dat.plot %>% 
  filter(group == "MSA",
         score > 0.39) %>% 
  mutate(lrst = paste0(ligand, receptor, source, target))

dat.pd <- dat.plot %>% 
  mutate(lrst = paste0(ligand, receptor, source, target)) %>% 
  filter(group == "PD",
         lrst %in% dat.msa$lrst)

dat.rel <- dat.msa %>% 
  mutate(score = score - dat.pd$score)

dat.rel %>% 
  lianaCircos()
```

# Figure 5 & EDF7

## Load data

```{r}
cao <- qread("cao_micro_pvm.qs", nthreads = 10)
cao_msa <- qread("cao_micro_pvm_msa.qs", nthreads = 10)
cao_pd <- qread("cao_micro_pvm_pd.qs", nthreads = 10)
cao_dis <- qread("cao_micro_pvm_dis.qs", nthreads = 10)
```

```{r}
sample.groups <- cao$sample.groups
```

## Figure 5a

```{r}
cao$plotEmbedding(groups = anno.micro,
                  size = 0.5, 
                  palette = pal.major, 
                  font.size = 3, 
                  raster = T, 
                  mark.groups = T,
                  plot.na = F,
                  alpha = 0.2) + 
  labs(x="UMAP1", y= "UMAP2") +
  theme(line = element_blank())
```

## Figure 5b

```{r, fig.width=4, fig.height=4}
cm.merged <- cao$data.object$getJointCountMatrix()

markers <- c("AIF1","F13A1","MRC1","CD163","CD74")

dotPlot(markers, 
        cm.merged, 
        cao$cell.groups, 
        cols = c("white","firebrick"))
```

## Figure 5d-f

```{r}
anno.sort <- anno.micro[rownames(cao$data.object$embedding)]
anno.sel <- anno.sort[!anno.sort %in% c("MIC_intermediate1", "PVMs")] %>% factor()
anno.sel2 <- anno.sort[!anno.sort %in% c("MIC_intermediate2", "MIC_activated")] %>% factor()

ldata1 <- getTscanTrajectory(cao$data.object, anno.sel)
ldata2 <- getTscanTrajectory(cao$data.object, anno.sel2)

ldata <- rbind(ldata1, ldata2)
```

### Figure 5d

```{r}
cao$data.object$embedding %>% 
  as.data.frame() %>% 
  mutate(., unname(anno.micro[rownames(.)])) %>% 
  setNames(c("UMAP1", "UMAP2", "annotation")) %>% 
  ggplot() + 
  geom_point(aes(UMAP1, UMAP2, col = annotation), size = 0.3) +
  geom_line(data = ldata, mapping=aes(UMAP1, UMAP2, group = edge), linewidth = 1) +
  theme_bw() + 
  theme(legend.position = "right",
        line = element_blank()) + 
  labs(col = "", x = "UMAP1", y = "UMAP2") +
  scale_colour_manual(values = pal.major) +
  dotSize(3)
```

## Figures 5e,f & EDF7

Prepare data

```{r}
emb <- cao$data.object$embedding %>% 
  `colnames<-`(c("UMAP1","UMAP2")) %>% 
  .[rownames(.) %in% names(anno.sel2), ]

sds_obj <- slingshot(emb, 
                     anno.sel2, 
                     start.clus = "MIC_steady-state",
                     stretch = 0
)

sds <- as.SlingshotDataSet(sds_obj)

pseudotime <- sds_obj@assays@data@listData$pseudotime[, 1]
```

Here, we show the code for running the generalized additive mixed model. It takes a significant amount of time to run, we provide data in `pseudotime_micro_l2.qs`.

```{r, eval = F}
mat <- cao$data.object$getJointCountMatrix() %>%
  .[match(names(sds_obj), rownames(.)), colSums(.) > 2e2] %>% 
  .[, !grepl(pattern = "RPL|RPS|MT-", colnames(.))]

mt <- mat %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column() %>%
  {message("Mutating"); mutate(.,
                               pseudotime = pseudotime[rowname],
                               condition = cao$data.object$getDatasetPerCell()[rowname] %>%
                                 as.character() %>%
                                 strsplit("_") %>%
                                 sget(1),
                               sample = strsplit(rowname, "!!") %>%
                                 sget(1))} %>%
  .[complete.cases(.), ] %>%
  {message("Melting"); reshape2::melt(., id.vars = c("rowname", "pseudotime", "condition", "sample"))} %$%
  {message("Splitting"); split(., variable)}
```

```{r, eval = F}
library(gamm4)

res <- mt %>% 
  sccore::plapply(\(x) {
    fit_full <- gamm4(data = x, formula = value ~ s(pseudotime, by = factor(condition)), random = ~(1 | sample), REML = F)
    fit_reduced <- gamm4(data = x, formula = value ~ s(pseudotime), random = ~(1 | sample), REML = F)
    fit_none <- gamm4(data = x, formula = value ~ 1, random = ~(1 | sample), REML = F)

    ann <- anova(fit_full$mer, fit_reduced$mer, fit_none$mer)

    if (any(ann$`Pr(>Chisq)` <= 0.05, na.rm = T)) {
      residuals <- predict(fit_full$gam, se.fit = T)$fit %>% 
        unname()

      r.sq <- c(summary(fit_full$gam)$r.sq, summary(fit_reduced$gam)$r.sq) %>% 
        setNames(c("full", "reduced"))

      out <- list(anova = ann,
                  residuals = residuals,
                  r.sq = r.sq)

      return(out)
    }
  }, n.cores = 40, mc.preschedule = T, mc.cleanup = T, progress = F) %>%
  .[!sapply(., is.null)]

qsave(res, "pseudotime_micro_l2.qs")
```

### Figure 5e

```{r}
cao$data.object$embedding %>% 
  as.data.frame() %>% 
  mutate(., unname(anno.micro[rownames(.)])) %>% 
  setNames(c("UMAP1", "UMAP2", "annotation")) %>% 
  mutate(., pseudotime = pseudotime[rownames(.)]) %>%
  ggplot() +
  geom_point(aes(UMAP1, UMAP2, col = pseudotime), size = 0.3) +
  geom_line(data = ldata2, aes(UMAP1, UMAP2, group = edge), size = 1) +
  theme_bw() +
  theme(line = element_blank()) +
  scale_color_gradient(low = "navyblue", high = "orange") +
  labs(col = "Pseudotime", x = "UMAP1", y = "UMAP2")
```

### Figure 5f

Please note, row order may change per iteration.

```{r, fig.height=4, fig.width=4}
res <- qread("pseudotime_micro_l2.qs")

rsq.pseudo <- res %>% 
  sapply(\(gene) gene$r.sq[1]) %>% 
  sort(decreasing = T)

res.filter <- res[names(rsq.pseudo)[seq(50)] %>% strsplit(".", fixed = T) %>% sget(1)]

cm.merged <- cao$data.object$getJointCountMatrix() %>% 
  .[match(names(pseudotime), rownames(.)), colnames(.) %in% names(res.filter)] %>% 
  .[rowSums(.) > 0, ]

## Predict smoothend expression 
pseudotime.both <- pseudotime[rownames(cm.merged)]
weights.both <- sds_obj@assays@data@listData$weights %>% .[match(names(pseudotime.both), rownames(.)), 1] + 1E-7

scFit <- cm.merged %>% 
  Matrix::t() %>% 
  tradeSeq::fitGAM(pseudotime = pseudotime.both, cellWeights = weights.both, verbose = T)

Smooth <- tradeSeq::predictSmooth(scFit, gene = colnames(cm.merged), tidy = F, n=100)

# Average across replicates and scale
Smooth <- t(scale(t(Smooth)))

# Seriate the results
Smooth <- Smooth[ seriation::get_order(seriation::seriate(Smooth, method="PCA_angle")), ]

# Create heatmap
col_fun = circlize::colorRamp2(c(-4, 0, 4), c("navy", "white", "firebrick"))

Heatmap(Smooth, 
        col = col_fun,
        cluster_columns=F, 
        cluster_rows=F, 
        show_column_names = F, 
        row_names_gp = grid::gpar(fontsize = 5))
```

### Extended Data Figure 7

We provide this figure here since all data are already loaded.

```{r, fig.height=40, fig.width=20}
cpc <- pseudotime %>% 
  names() %>% 
  strsplit("_") %>% 
  sget(1) %>% 
  factor()

res[rownames(Smooth)] %>% 
  lget("residuals") %>% 
  lapply(as.data.frame) %>% 
  lapply(setNames, "residuals") %>% 
  lapply(mutate, group = cpc, pseudotime = pseudotime) %>% 
  data.table::rbindlist(idcol = "gene") %>% 
  ggplot(aes(pseudotime, residuals, col = group)) +
  geom_smooth() +
  theme_bw() +
  facet_wrap(~gene, ncol = 5, scales = "free")
```

## Figure 5g

```{r, fig.height=4, fig.width=3}
cao_msa$plotCellLoadings(show.pvals = F, 
                         alpha = 0)$data %>% 
  ggplot(aes(ind, values, fill = ind)) +
  geom_hline(yintercept = 0, col = "black") +
  geom_violin() + 
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = pal.major) + 
  theme(legend.position = "none") +
  scale_y_continuous(breaks = c(-1,0,1), 
                     limits = c(-1,1),
                     labels = c("-1\nCTRL",0,"1\nMSA")) +
  geom_vline(xintercept = 4.5, col = "red") +
  labs(x = "", y = "") +
  theme(line = element_blank())
```

## Figure 5h

```{r, fig.height=4, fig.width=3}
cao_pd$plotCellLoadings(show.pvals = F, 
                        alpha = 0)$data %>% 
  ggplot(aes(ind, values, fill = ind)) + 
  geom_hline(yintercept = 0, col = "black") +
  geom_violin() + 
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = pal.major) + 
  theme(legend.position = "none") +
  scale_y_continuous(breaks = c(-1,0,1), 
                     limits = c(-1,1),
                     labels = c("-1\nCTRL",0,"1\nPD")) +
  geom_vline(xintercept = 4.5, col = "red") +
  labs(x = "", y = "") +
  theme(line = element_blank())
```

## Figure 5i

```{r, fig.height=4, fig.width=3}
cao_dis$plotCellLoadings(show.pvals = F, 
                         alpha = 0)$data %>% 
  ggplot(aes(ind, values, fill = ind)) + 
  geom_hline(yintercept = 0, col = "black") +
  geom_violin() + 
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = pal.major) + 
  theme(legend.position = "none") +
  scale_y_continuous(breaks = c(-1,0,1), 
                     limits = c(-1,1),
                     labels = c("-1\nPD",0,"1\nMSA")) +
  geom_vline(xintercept = 3.5, col = "red") +
  labs(x = "", y = "") +
  theme(line = element_blank())
```

## Figure 5k

We provide combined data from the RNAscope experiments as a single file `res.qs`.

```{r}
res <- qread("RNAscope.qs") %>% 
  filter(target == "AIF1", Ch1NumSpots > 0) %>% 
  arrange(file)

area <- read.table("RNAscope_areas.tsv", sep = "\t", header = T) %>% # Million pixels
  mutate(file = paste(File.name, Tissue, sep = "")) %>% 
  filter(file %in% res$file) %>% 
  arrange(file) %>% 
  mutate(rel_px = px.2 / 1E6)
```

```{r, fig.width=4, fig.height=8}
p1 <- res %>% 
  group_by(group, target, file) %>% 
  summarize(spots = mean(Ch1NumSpots)) %>% 
  ggplot(aes(group, spots, fill = group)) + 
  geom_boxplot() +
  geom_jitter(width = 0.2) +
  theme_bw() +
  labs(y = "Mean no. spots per double-positive cell", x = "") +
  stat_compare_means(method = "kruskal.test", label.y = 115) +
  stat_compare_means(comparisons = comp, method = "wilcox.test") +
  theme(line = element_blank()) +
  guides(fill = "none") +
  scale_fill_manual(values = pal.major)

p2 <- res %>% 
  group_by(group, target, file) %>% 
  summarize(no_cells = n()) %>% 
  as.data.frame() %>% 
  arrange(file) %>% 
  mutate(rel_cells = no_cells / area$rel_px) %>% 
  ggplot(aes(group, rel_cells, fill = group)) + 
  geom_boxplot(outliers = F) +
  geom_jitter(width = 0.2) +
  theme_bw() +
  stat_compare_means(method = "kruskal.test", label.y = 17) +
  stat_compare_means(comparisons = comp, method = "wilcox.test") +
  labs(y = "No. double-positive cells\nNormalized to area", x = "") +
  theme(line = element_blank()) +
  guides(fill = "none") +
  scale_fill_manual(values = pal.major)

plot_grid(plotlist = list(p1, p2), ncol = 1)
```

## Figure 5l

```{r, fig.width=12, fig.height=3}
fams <- cao_msa$test.results[["GSEA"]]$families

df.all <- list(cao_msa$.__enclos_env__$private$getOntologyPvalueResults(name = "GSEA", genes = "up", p.adj = 1, q.value = 1),
               cao_msa$.__enclos_env__$private$getOntologyPvalueResults(name = "GSEA", genes = "down", p.adj = 1, q.value = 1)) %>% 
  bind_rows() %>% 
  mutate(logp = -log10(p.adjust))

# Relevant pathways
p.ids <- c("GO:0060760", "GO:0002230", "GO:0019915", "GO:1904646", "GO:0042775", "GO:0019646", "GO:0033108", "GO:0035455", "GO:0061077", "GO:0006914", "GO:0042026", "GO:0034340", "GO:0051607", "GO:0060337", "GO:0035455", "GO:0002684", "GO:0036037", "GO:0048514")

df <- cacoa:::getOntologyFamilyChildren(df.all, fams=fams, type = "GSEA", subtype = "BP") %>% 
  mutate(Group = paste0("MIC_", tolower(Group)) %>% gsub("MIC_pvms", "PVMs", .)) %>% 
  filter(ID %in% p.ids)

df.all %<>% mutate(Group = paste0("MIC_", tolower(Group)) %>% gsub("MIC_pvms", "PVMs", .))

p1 <- ggplot(df.all, aes(NES, logp, col = Group, shape = Group)) +
  geom_point(data = ~filter(., p.adjust > 0.05), size = 0.1, alpha = 0.5, col = "black") +
  geom_point(data = ~filter(., p.adjust <= 0.05), size = 2, alpha = 0.5) +
  theme_bw() +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  scale_color_manual(values = pal.major) +
  theme(line = element_blank()) +
  labs(y = "-log10(adj. p)") +
  dotSize(3)

p2 <- df %>% 
  filter(p.adjust <= 0.05, NES > 0) %>% 
  mutate(Group = factor(Group, levels = sort(unique(Group)))) %>% 
  select(Description, NES, Group) %$% 
  split(., Group) %>% 
  lapply(dplyr::slice, seq(5)) %>% 
  bind_rows() %>% 
  dplyr::slice(seq(25)) %>% 
  mutate(., x = 0, y = rev(seq(nrow(.)))) %>% 
  ggplot(aes(x, y, label = Description, col = Group)) + 
  geom_text(hjust = 0) +
  theme_void() +
  xlim(0, 1) +
  labs(title = "Selected terms for up-regulated genes") +
  scale_color_manual(values = pal.major) +
  guides(col = "none")

p3 <- df %>% 
  filter(p.adjust <= 0.05, NES < 0) %>% 
  select(Description, NES, Group) %>%
  dplyr::slice(seq(20)) %>% 
  mutate(., x = 0, y = rev(seq(nrow(.)))) %>% 
  ggplot(aes(x, y, label = Description, col = Group)) + 
  geom_text(hjust = 0) +
  theme_void() +
  xlim(0, 1) +
  labs(title = "Selected terms for down-regulated genes") +
  scale_color_manual(values = pal.major) +
  guides(col = "none")

plot_grid(plotlist = list(p3, p1, p2), ncol = 3, rel_widths = c(0.9, 1, 0.9))
```

## Figure 5m

```{r, fig.width=12, fig.height=3}
fams <- cao_pd$test.results[["GSEA"]]$families

df.all <- list(cao_pd$.__enclos_env__$private$getOntologyPvalueResults(name = "GSEA", genes = "up", p.adj = 1, q.value = 1),
               cao_pd$.__enclos_env__$private$getOntologyPvalueResults(name = "GSEA", genes = "down", p.adj = 1, q.value = 1)) %>% 
  bind_rows() %>% 
  mutate(logp = -log10(p.adjust))

# Select relevant pathways
p.ids <- c("GO:0042776", "GO:0034341", "GO:0001916", "GO:0002503", "GO:0019885", "GO:0016042", "GO:0001909", "GO:0051085", "GO:0045824", "GO:0044000", "GO:0019885", "GO:0044242", "GO:0001944", "GO:0034976", "GO:0019646", "GO:0009205", "GO:0019883", "GO:0070972", "GO:0002474")

df <- cacoa:::getOntologyFamilyChildren(df.all, fams=fams, type = "GSEA", subtype = "BP") %>% 
  mutate(Group = paste0("MIC_", tolower(Group)) %>% gsub("MIC_pvms", "PVMs", .)) %>% 
  filter(ID %in% p.ids)

df.all %<>% mutate(Group = paste0("MIC_", tolower(Group)) %>% gsub("MIC_pvms", "PVMs", .))

p1 <- ggplot(df.all, aes(NES, logp, col = Group, shape = Group)) +
  geom_point(data = ~filter(., p.adjust > 0.05), size = 0.1, alpha = 0.5, col = "black") +
  geom_point(data = ~filter(., p.adjust <= 0.05), size = 2, alpha = 0.5) +
  theme_bw() +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  scale_color_manual(values = pal.major) +
  theme(line = element_blank()) +
  labs(y = "-log10(adj. p)") +
  dotSize(3)

p2 <- df %>% 
  filter(p.adjust <= 0.05, NES > 0) %>% 
  select(Description, NES, Group) %$% 
  split(., Group) %>% 
  lapply(dplyr::slice, seq(5)) %>% 
  bind_rows() %>% 
  dplyr::slice(seq(25)) %>% 
  mutate(., x = 0, y = rev(seq(nrow(.)))) %>% 
  ggplot(aes(x, y, label = Description, col = Group)) + 
  geom_text(hjust = 0) +
  theme_void() +
  xlim(0, 1) +
  labs(title = "Selected terms for up-regulated genes") +
  scale_color_manual(values = pal.major) +
  guides(col = "none")

p3 <- df %>% 
  filter(p.adjust <= 0.05, NES < 0) %>% 
  select(Description, NES, Group) %$% 
  split(., Group) %>% 
  lapply(dplyr::slice, seq(5)) %>% 
  bind_rows() %>% 
  dplyr::slice(seq(25)) %>% 
  mutate(., x = 0, y = rev(seq(nrow(.)))) %>% 
  ggplot(aes(x, y, label = Description, col = Group)) + 
  geom_text(hjust = 0) +
  theme_void() +
  xlim(0, 1) +
  labs(title = "Selected terms for down-regulated genes") +
  scale_color_manual(values = pal.major) +
  guides(col = "none")

plot_grid(plotlist = list(p3, p1, p2), ncol = 3, rel_widths = c(0.9, 1, 0.9))
```

# Figure 6

Figures 6c,e,f where made with GraphPad Prism. Data are available in `Phagocytosis_assay_triplicates_raw_data_HH.tsv` (Copenhagen experiment) or `HOUSTON.tsv` (Houston experiment).

## Figure 6b

```{r, fig.height=4, fig.width=5}
dat <- read.delim("Phagocytosis_assay_triplicates_raw_data_HH.tsv", 
                  header = T, dec = ",") %>% 
  tidyr::pivot_longer(names_to = "group", cols = -1, values_to = "value") %>% 
  mutate(group = gsub(".1|.2", "", group) %>% factor(labels = c("PBS","LPS","MSA CSF","CTRL CSF","PD CSF"))) %>% 
  mutate(group = factor(group, levels = c("PBS","LPS","CTRL CSF","MSA CSF","PD CSF"))) %>% 
  group_by(Time, group) %>% 
  summarize(mean = mean(value, na.rm = T), 
            sd = sd(value, na.rm = T))

dat %>% 
  ggplot(aes(Time, mean, group=group, color=group)) + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +
  geom_point(size = 0.5) +
  theme_bw() +
  labs(x = "Time (hours)", y = "pHrodo-labelled E. coli uptake - intensity ratio [AU]", col = "Stimulation") +
  scale_color_manual(values = pal.major) +
  theme(line = element_blank())
```

## Figure 6d

```{r, fig.width=4, fig.height=4}
dat.raw <- read.table("Cytokine_summary.tsv", header = TRUE, sep = "\t", dec = ",")

dat.tmp <- dat.raw %>%
  melt(id.vars = c("Sample","Group","Condition")) %>% 
  mutate(variable = variable %>% 
           as.character() %>% 
           gsub(".", "-", ., fixed = T) %>% 
           as.factor()) %>%
  filter(Condition == "Media", !Group %in% c("PBS","LPS")) %>%
  filter(Condition == "Media") %>%
  mutate(Group = Group %>% factor(levels = c("CTRL","MSA","PD")),
         variable = variable %>% factor())

dat.stat <- dat.raw %>% 
  select(!IL.8) %>% 
  mutate(Group = factor(Group, levels = c("PBS","LPS","CTRL","MSA","PD"))) %>% 
  filter(!Group %in% c("PBS","LPS"))

res.il10 <- FSA::dunnTest(IL10 ~ Group, data = dat.stat %>% {`colnames<-`(., gsub(".", "", colnames(.), fixed = T))}, method = "bh")

stat.test <- dat.tmp %>% 
  filter(variable == "IL-10") %>% 
  select(Group, value) %>% 
  as.data.frame() %>%
  rstatix::wilcox_test(value ~ Group) %>% 
  mutate(p = res.il10$res$P.unadj,
         p.adj = res.il10$res$P.adj %>% formatC(digits = 2),
         p.adj.signif = c("*","ns","ns")) %>% 
  rstatix::add_xy_position(x = "Group")

dat.tmp %>% 
  filter(variable == "IL-10") %>%
  ggplot(aes(Group, value)) + 
  geom_point(aes(col = Group), size = 3, position = position_jitter(width = 0.3)) + 
  theme_bw() +
  theme(line = element_blank()) +
  stat_pvalue_manual(stat.test, label = "p.adj", tip.length = 0.01, backet.nudge.y = 3, hide.ns = F) +
  scale_color_manual(values = pal.major) +
  labs(x = "", y = "IL-10 pg/mL") +
  guides(col = "none")
```

## Figures 6g-i

```{r}
dat.melt <- read.table("Microglia+CSF_samples_aSyn_sCD163.tsv", header = T, sep = "\t", dec = ",") %>%  
  melt(id.vars = c("diagnosis", "sex")) %>% 
  mutate(diagnosis = factor(diagnosis, levels = c("CTRL","MSA","IPD"), labels = c("CTRL","MSA","PD")))
```

### Figure 6g

```{r, fig.width=4, fig.height=4}
dat.melt %>% 
  filter(variable == "sCD163_ng.mL_CSF") %>% 
  ggplot(aes(diagnosis, value)) +
  geom_boxplot(aes(fill = diagnosis)) +
  theme_bw() + 
  labs(x = "", y = "sCD163 ng/ml", fill = "Diagnosis") +
  scale_fill_manual(values = pal.major) +
  stat_compare_means(method = "t.test", comparisons = list(c("CTRL","MSA"), c("MSA","PD"), c("CTRL","PD"))) +
  guides(fill = "none") +
  theme(line = element_blank())
```

### Figure 6h

```{r, fig.width=4, fig.height=4}
dat.melt %>% 
  filter(variable == "aSyn_pg.mL") %>% 
  ggplot(aes(diagnosis, value)) +
  geom_boxplot(aes(fill = diagnosis)) +
  theme_bw() + 
  labs(x = "", y = "aSyn pg/ml", fill = "Diagnosis") +
  scale_fill_manual(values = pal.major) +
  stat_compare_means(method = "t.test", comparisons = list(c("CTRL","MSA"), c("MSA","PD"), c("CTRL","PD"))) +
  guides(fill = "none") +
  theme(line = element_blank())
```

### Figure 6i

```{r, fig.width=4, fig.height=4}
dat.melt %>% 
  filter(variable %in% c("aSyn_pg.mL","sCD163_ng.mL_CSF")) %>%
  select(-sex) %>% 
  mutate(id = rep(seq(63), 2)) %>% 
  tidyr::pivot_wider(names_from = variable, values_from = value) %>% 
  ggplot(aes(aSyn_pg.mL, sCD163_ng.mL_CSF)) +
  geom_point(aes(col = diagnosis)) +
  theme_bw() +
  labs(x = "aSyn pg/ml", y = "sCD163, ng/ml", col = "") +
  geom_smooth(method = MASS::rlm, se = FALSE, color = "black") +
  stat_poly_eq(aes(label = paste(after_stat(rr.label), after_stat(p.value.label), sep = "*\", \"*")), color = "black") +
  scale_color_manual(values = pal.major) +
  theme(line = element_blank())
```

# Extended Data Figure 1

## Load data

```{r}
samples <- anno.major %>% 
  names() %>% 
  strsplit("!!") %>% 
  sget(1) %>% 
  unique()

crm <- qread("crm.qs",
             nthreads = 10)
```

## 1a-f

```{r, fig.width=8, fig.height=12}
mtp <- crm$selectMetrics(ids = c(1:4,6,19))

mtp %>% 
  lapply(\(met) crm$plotSummaryMetrics(metrics = met, comp.group = "sample", second.comp.group = "group") + 
           scale_fill_manual(values = pal.major) +
           labs(x = "") + 
           theme(axis.text.x = element_blank(),
                 axis.ticks.x = element_blank())) %>% 
  plot_grid(plotlist = ., labels = letters[1:6], ncol = 2)
```

## 1g

```{r, fig.width=6, fig.height=4}
crm$plotSummaryMetrics(metrics = crm$selectMetrics(ids = 20), comp.group = "sample", second.comp.group = "group") + 
  scale_fill_manual(values = pal.major) +
  labs(x = "") + 
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "right")
```

## 1h

```{r, fig.width=6, fig.height=4}
crm$plotFilteredCells(doublet.method = "doubletdetection", depth.cutoff = 5e2, size = 0.2, alpha = 0.1) +
  dotSize(3)
```

## 1i

```{r, fig.width=8, fig.height=4}
crm$plotFilteredCells(type = "bar", doublet.method = "doubletdetection", depth.cutoff = 5e2)$data %>%  
  filter(sample != "MSA_1406") %>%
  mutate(sample = strsplit(sample, "_") %>% 
           sget(1)) %$% 
  split(., sample) %>% 
  lapply(\(x) split(x, x$filter)) %>% 
  lapply(lapply, \(df) mutate(df, sample = paste0(sample, seq(nrow(df))))) %>% 
  lapply(bind_rows) %>% 
  bind_rows() %>%
  mutate(sample = factor(sample, levels = c(paste0("CTRL", seq(10)), paste0("MSA", seq(7)), paste0("PD", seq(12))))) %>% 
  ggplot(aes(sample, pct, fill = filter)) + 
  geom_bar(stat = "identity") + 
  geom_text_repel(aes(label = sprintf("%0.2f", round(pct, digits = 2))), 
                  position = position_stack(vjust = 0.5), 
                  direction = "y", 
                  size = 2.5) + 
  crm$theme + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "", y = "Percentage cells filtered") +
  theme(legend.position = "bottom")
```

# Extended Data Figure 2

These figures cannot be reproduced here due to GDPR. Leaving the code for visibility.

## Load data

```{r, eval = F}
crm <- qread("crm.qs", nthreads = 10)
```

## 2a

```{r, fig.width=8, fig.height=8, eval = F}
crm$plotCbCells(samples = crm$metadata$sample %>% as.character() %>% .[. != "MSA_1406"]) +
  scale_x_discrete(labels = c(paste("CTRL", seq(10)), paste("MSA", seq(7)), paste("PD", seq(12)))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

## 2b

```{r, fig.width=5, fig.height=4, eval = F}
crm$plotCbAmbGenes() +
  theme(line = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.ticks.x = element_line())
```

# Extended Data Figure 3

## Load and prepare data

```{r}
con$embedding <- con$embeddings$UMAP

sample.groups <- con$samples %>% 
  names() %>% 
  setNames(grepl.replace(., c("CTRL","MSA","PD")), .)

cao <- Cacoa$new(data.object = con, 
                 sample.groups = ifelse(grepl("CTRL", sample.groups), "CTRL", "DISEASE") %>% 
                   setNames(names(sample.groups)), 
                 cell.groups = anno.major, 
                 ref.level = "CTRL", 
                 target.level = "DISEASE", 
                 n.cores = 32)

meta <- read.delim("metadata.tsv", sep = "\t") %>% 
  filter(sample %in% names(con$samples)) %>% 
  tibble::column_to_rownames("sample")

cao$estimateExpressionShiftMagnitudes()
cao$estimateMetadataSeparation(sample.meta = meta, 
                               space = "expression.shifts")
```

## 3a

```{r, fig.width = 6, fig.height=4}
cao$plotSampleDistances(space = "expression.shifts", show.sample.size = T, method = "UMAP")$data %>% 
  mutate(condition = unname(sample.groups)) %>% 
  ggplot(aes(x, y, shape = condition, col = condition)) +
  geom_point(aes(size = n.cells)) + scale_size_continuous(trans = "log10", 
                                                          range = c(5 * 0.5, 5 * 1.5), name = "Num. cells") +
  theme_bw() +
  scale_color_manual(values = pal.major) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(x = "", y = "", title = "Condition", shape = "Condition", col = "Condition") +
  dotSize(3)
```

## 3b

```{r, fig.width = 6, fig.height=4}
cao$plotSampleDistances(space = "expression.shifts", method = "UMAP", sample.colors = meta$age %>% setNames(rownames(meta)))$data %>% 
  mutate(condition = unname(sample.groups)) %>% 
  ggplot(aes(x, y, shape = condition, col = color)) +
  geom_point(size = 3) +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(x = "", y = "", title = "Age", color = "Age (y)", shape = "Condition")
```

## 3c

```{r, fig.width = 6, fig.height=4}
cao$plotSampleDistances(space = "expression.shifts", method = "UMAP", sample.colors = meta$brain_bank %>% setNames(rownames(meta)))$data %>% 
  mutate(condition = unname(sample.groups)) %>% 
  ggplot(aes(x, y, shape = condition, col = color)) +
  geom_point(size = 3) +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(x = "", y = "", title = "Brain bank", shape = "Condition", color = "Brain bank")
```

## 3d

```{r, fig.width = 6, fig.height=4}
cao$plotSampleDistances(space = "expression.shifts", method = "UMAP", sample.colors = meta$extract %>% setNames(rownames(meta)))$data %>% 
  mutate(condition = unname(sample.groups)) %>% 
  ggplot(aes(x, y, shape = condition, col = color)) +
  geom_point(size = 3) +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(x = "", y = "", title = "Extract", shape = "Condition", color = "Extract")
```

## 3d

```{r, fig.width = 6, fig.height=4}
cao$plotSampleDistances(space = "expression.shifts", method = "UMAP", sample.colors = meta$flowcell %>% setNames(rownames(meta)))$data %>% 
  mutate(condition = unname(sample.groups)) %>% 
  ggplot(aes(x, y, shape = condition, col = color)) +
  geom_point(size = 3) +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(x = "", y = "", title = "Flowcell", shape = "Condition", color = "Flowcell")
```

# Extended Data Figure 5

```{r}
cm.merged <- con$getJointCountMatrix(raw = T) %>% 
  Matrix::t() %>% 
  .[, colnames(.) %in% names(anno.major)]

# Create sample-wise annotation
anno.donor <- con$getDatasetPerCell()[colnames(cm.merged)]
anno.subtype <- anno.major %>% 
  .[!is.na(.)] %>% 
  factor()

idx <- intersect(anno.donor %>% names(), anno.subtype %>% names())

anno.donor %<>% .[idx]
anno.subtype %<>% .[idx] %>% 
  {`names<-`(as.character(.), names(.))}

anno.final <- paste0(anno.donor,"!!",anno.subtype) %>% 
  `names<-`(anno.donor %>% names())

# Create pseudo CM
cm.pseudo <- sccore::collapseCellsByType(cm.merged %>% Matrix::t(), 
                                         groups = anno.final, min.cell.count = 1) %>% 
  t() %>%
  apply(2, as, "integer")

cm.pseudo %<>% 
  {. + 1} %>% 
  DESeq2::DESeqDataSetFromMatrix(., 
                                 colnames(.) %>% 
                                   strsplit("_") %>% 
                                   sget(1) %>% 
                                   data.frame() %>% 
                                   `dimnames<-`(list(colnames(cm.pseudo), "group")), 
                                 design = ~ group) %>% 
  DESeq2::estimateSizeFactors() %>% 
  DESeq2::counts(normalized = T)
```

```{r, fig.width=12, fig.height=24}
c("TSPO", "COQ2", "MAPT", "FBXO47", "ELOVL7", "EDN1", "GAB1", "TENM2", "RABGEF1", "PLA2G4C", "INPP4B", "ZIC1", "ZIC2", "ZIC3", "ZIC4", "SNCA", "TPPP", "TPPP") %>% 
  {Map(\(gene, leg) plotGenePseudoBulk(gene, cm.pseudo, leg), gene = ., leg = c(rep(F, 17), T))} %>% 
  cowplot::plot_grid(plotlist = ., ncol = 3)
```

# Extended Data Figures 6 & 8

## Load and prepare data

```{r, fig.width=6, fig.height=4}
cao <- qread("cao_micro_pvm.qs", nthreads = 10)

anno.sort <- anno.micro[rownames(cao$data.object$embedding)]
anno.sel <- anno.sort[!anno.sort %in% c("MIC_intermediate1", "PVMs")] %>% factor()

emb <- cao$data.object$embedding %>% 
  `colnames<-`(c("UMAP1","UMAP2")) %>% 
  .[rownames(.) %in% names(anno.sel), ]

sds_obj <- slingshot(emb, 
                     anno.sel, 
                     start.clus = "MIC_steady-state",
                     stretch = 0
)

sds <- as.SlingshotDataSet(sds_obj)

pseudotime <- sds_obj@assays@data@listData$pseudotime[, 1]
```

## 6a

```{r, fig.width=6, fig.height=4}
plot.df <- cao$data.object$embedding %>% 
  as.data.frame() %>% 
  .[rownames(.) %in% names(sds_obj), ] %>% 
  `colnames<-`(c("UMAP1","UMAP2")) %>% 
  mutate(., annotation = anno.micro[rownames(.)] %>% as.factor()) %>% 
  .[complete.cases(.),]

ldata <- getTscanTrajectory(cao$data.object, anno.sel)

cao$data.object$embedding %>% 
  as.data.frame() %>% 
  mutate(., unname(anno.micro[rownames(.)])) %>% 
  setNames(c("UMAP1", "UMAP2", "annotation")) %>% 
  mutate(., pseudotime = pseudotime[rownames(.)]) %>%
  ggplot() +
  geom_point(aes(UMAP1, UMAP2, col = pseudotime), size = 0.3) +
  geom_line(data = ldata1, aes(UMAP1, UMAP2, group = edge), size = 1) +
  theme_bw() +
  theme(line = element_blank()) +
  scale_color_gradient(low = "navyblue", high = "orange") +
  labs(col = "Pseudotime", x = "UMAP1", y = "UMAP2")# +
```

## 6b

Here, we show the code for running the generalized additive mixed model. It takes a significant amount of time to run, we provide data in `pseudotime_micro_l1.qs`.

```{r, eval = F}
mat <- cao$data.object$getJointCountMatrix() %>%
  .[match(names(sds_obj), rownames(.)), colSums(.) > 2e2] %>% 
  .[, !grepl(pattern = "RPL|RPS|MT-", colnames(.))]

mt <- mat %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column() %>%
  {message("Mutating"); mutate(.,
                               pseudotime = pseudotime[rowname],
                               condition = cao$data.object$getDatasetPerCell()[rowname] %>%
                                 as.character() %>%
                                 strsplit("_") %>%
                                 sget(1),
                               sample = strsplit(rowname, "!!") %>%
                                 sget(1))} %>%
  .[complete.cases(.), ] %>%
  {message("Melting"); reshape2::melt(., id.vars = c("rowname", "pseudotime", "condition", "sample"))} %$%
  {message("Splitting"); split(., variable)}
```

```{r, eval = F}
library(gamm4)

res <- mt %>% 
  sccore::plapply(\(x) {
    fit_full <- gamm4(data = x, formula = value ~ s(pseudotime, by = factor(condition)), random = ~(1 | sample), REML = F)
    fit_reduced <- gamm4(data = x, formula = value ~ s(pseudotime), random = ~(1 | sample), REML = F)
    fit_none <- gamm4(data = x, formula = value ~ 1, random = ~(1 | sample), REML = F)

    ann <- anova(fit_full$mer, fit_reduced$mer, fit_none$mer)

    if (any(ann$`Pr(>Chisq)` <= 0.05, na.rm = T)) {
      residuals <- predict(fit_full$gam, se.fit = T)$fit %>% 
        unname()

      r.sq <- c(summary(fit_full$gam)$r.sq, summary(fit_reduced$gam)$r.sq) %>% 
        setNames(c("full", "reduced"))

      out <- list(anova = ann,
                  residuals = residuals,
                  r.sq = r.sq)

      return(out)
    }
  }, n.cores = 40, mc.preschedule = T, mc.cleanup = T, progress = F) %>%
  .[!sapply(., is.null)]

qsave(res, "pseudotime_micro_l1.qs")
```

```{r, fig.height=4, fig.width=6}
res <- qread("pseudotime_micro_l1.qs")

rsq.pseudo <- res %>% 
  sapply(\(gene) gene$r.sq[1]) %>% 
  sort(decreasing = T)

res.filter <- res[names(rsq.pseudo)[seq(50)] %>% strsplit(".", fixed = T) %>% sget(1)]

cm.merged <- cao$data.object$getJointCountMatrix() %>% 
  .[match(names(pseudotime), rownames(.)), colnames(.) %in% names(res.filter)] %>% 
  .[rowSums(.) > 0, ]

## Predict smoothend expression 
pseudotime.both <- pseudotime[rownames(cm.merged)]
weights.both <- sds_obj@assays@data@listData$weights %>% .[match(names(pseudotime.both), rownames(.)), 1] + 1E-7

scFit <- cm.merged %>% 
  Matrix::t() %>% 
  tradeSeq::fitGAM(pseudotime = pseudotime.both, cellWeights = weights.both, verbose = T)

Smooth <- tradeSeq::predictSmooth(scFit, gene = colnames(cm.merged), tidy = F, n=100)

# Average across replicates and scale
Smooth <- t(scale(t(Smooth)))

# Seriate the results
Smooth <- Smooth[ seriation::get_order(seriation::seriate(Smooth, method="PCA_angle")), ]

# Create heatmap
col_fun = circlize::colorRamp2(c(-4, 0, 4), c("navy", "white", "firebrick"))

Heatmap(Smooth, 
        col = col_fun,
        cluster_columns=F, 
        cluster_rows=F, 
        show_column_names = F, 
        row_names_gp = grid::gpar(fontsize = 5))
```

## 8

```{r, fig.height=40, fig.width=20}
cpc <- pseudotime %>% 
  names() %>% 
  strsplit("_") %>% 
  sget(1) %>% 
  factor()

res.filter %>% 
  lget("residuals") %>% 
  lapply(as.data.frame) %>% 
  lapply(setNames, "residuals") %>% 
  lapply(mutate, group = cpc, pseudotime = pseudotime) %>% 
  data.table::rbindlist(idcol = "gene") %>% 
  ggplot(aes(pseudotime, residuals, col = group)) +
  geom_smooth() +
  theme_bw() +
  facet_wrap(~gene, ncol = 5, scales = "free")
```

# Extended Data Figure 9

## 9a

We provide the results output from scHLAcount in `scHLAcount.zip`.

```{r}
samples <- dir("scHLAcount/rydbirk/") %>% .[grepl(pattern = "_", .)]

cms <- lapply(samples, function(x) {
  labels <- read.delim(paste0("scHLAcount/rydbirk/",x,"/labels.tsv"), header = F) %>% .[,1]
  barcodes <- read.delim(paste0("scHLAcount/rydbirk/",x,"/barcodes.tsv"), header = F) %>% 
    .[,1] %>%
    sapply(function(y) paste0(x,"one_",y))

  mtx <- Matrix::readMM(paste0("scHLAcount/rydbirk/",x,"/count_matrix.mtx")) %>% 
    t() %>%
    `dimnames<-`(list(barcodes, labels)) %>%
    .[rowSums(.) > 0,] # Remove cells with no counts

  tmp <- colnames(mtx) %>%
    strsplit("*", T) %>%
    sapply(`[[`, 1)

  # If more than one allele per HLA type has been detected, sum per HLA type
  if(any(tmp %>% table() %>% as.numeric() > 1)) {
    cs <- colSums(mtx)
    alleles <- unique(tmp)

    mtx <- lapply(alleles, function(y) mtx[,tmp == y]) %>%
      lapply(function(y) {
        if(class(y) == "dgTMatrix") rowSums(y) else y
      }) %>% 
      unlist() %>% 
      Matrix(nrow = nrow(mtx)) %>%
      `rownames<-`(rownames(mtx))
  }

  colnames(mtx) <- unique(tmp)
  return(mtx)
}) %>%
  setNames(samples)
```

Calculations

```{r}
# Cell-wise
cell.counts <- sapply(cms, rowSums) %>% 
  unname() %>% 
  unlist() %>% 
  `names<-`(., gsub("one_","!!",names(.)))

depth <- getConosDepth(con) %>%
  .[match(names(cell.counts), names(.))] %>%
  {data.frame(cell = names(.),
              sample = names(.) %>% strsplit("!!", T) %>% sapply(`[[`, 1),
              depth = unname(.))}

depth[is.na(depth)] <- 0

cell.group <- grepl.replace(names(cell.counts), patterns = c("CTRL","PD","MSA")) %>% as.factor()

mhc1.raw <- sapply(cms, function(x) {
  tmp <- x[,colnames(x) %in% c("A","B","C")]
  if(class(tmp) != "numeric") rowSums(tmp) else tmp
}) %>% 
  unname() %>% 
  unlist() %>% 
  `names<-`(., gsub("one_","!!",names(.)))

mhc2.raw <- sapply(cms, function(x) {
  tmp <- x[,!colnames(x) %in% c("A","B","C")]
  if(class(tmp) != "numeric") rowSums(tmp) else tmp
}) %>% 
  unname() %>% 
  unlist() %>% 
  `names<-`(., gsub("one_","!!",names(.)))

cell.df <- data.frame(group = cell.group,
                      counts = cell.counts,
                      mhc1 = mhc1.raw,
                      mhc2 = mhc2.raw,
                      depth = depth$depth) %>%
  mutate(., anno = anno.major[match(rownames(.), names(anno.major))],
         sample = rownames(.) %>% sapply(strsplit, "!!", T) %>% sapply(`[[`, 1)) %>%
  mutate(type = grepl.replace(anno %>% as.character(), levels(anno), c("Brain","Brain","Brain","Peripheral","Peripheral","Peripheral","Brain","Brain","Brain","Brain"))) %>%
  `rownames<-`(names(cell.counts)) %>%
  filter(!is.na(anno)) 
```

Load Smajic data

```{r}
samples <- dir("scHLAcount/smajic") %>% .[grepl(pattern = "SRR", .)]

cms <- lapply(samples, function(x) {
  labels <- read.delim(paste0("scHLAcount/smajic/",x,"/labels.tsv"), header = F) %>% .[,1]
  barcodes <- read.delim(paste0("scHLAcount/smajic/",x,"/barcodes.tsv"), header = F) %>% 
    .[,1] %>%
    sapply(function(y) paste0(x,"_",y,"_1"))
  mtx <- Matrix::readMM(paste0("scHLAcount/smajic/",x,"/count_matrix.mtx")) %>% 
    t() %>%
    `dimnames<-`(list(barcodes, labels)) %>%
    .[rowSums(.) > 0,] # Remove cells with no counts

  tmp <- colnames(mtx) %>%
    strsplit("*", T) %>%
    sapply(`[[`, 1)

  # If more than one allele per HLA type has been detected, sum per HLA type
  if(any(tmp %>% table() %>% as.numeric() > 1)) {
    cs <- colSums(mtx)
    alleles <- unique(tmp)

    mtx <- lapply(alleles, function(y) mtx[,tmp == y]) %>%
      lapply(function(y) {
        if(class(y) == "dgTMatrix") rowSums(y) else y
      }) %>% 
      unlist() %>% 
      Matrix(nrow = nrow(mtx)) %>%
      `rownames<-`(rownames(mtx))
  }

  colnames(mtx) <- unique(tmp)
  return(mtx)
}) %>%
  setNames(samples)

# Correct naming
name.df <- data.frame(samples = samples, ids = c(rep("PD",6) %>% mapply(function(x,y) paste0(x,y), x = ., y = c(1,1,2:5)),
                                                 rep("CTRL",6) %>% mapply(function(x,y) paste0(x,y), x = ., y = 1:6)))

anno <- readRDS(paste0("scHLAcount/smajic/anno.sma.rds")) %>%
  setNames(., names(.) %>% 
             strsplit("_", T) %>% 
             lapply(function(x) {
               if(grepl("C", x[[1]])) x[[1]] %<>% gsub("C","CTRL", .)
               return(x)
             }) %>%
             sapply(function(x) paste0(x[[1]],"_",x[[2]])))

cms <- 1:12 %>% 
  lapply(function(x) {
    rnames <- cms[[x]] %>%
      rownames() %>%
      names() %>%
      sapply(function(y) paste0(name.df[x,2],"_",y)) %>%
      unname()

    rownames(cms[[x]]) <- rnames
    return(cms[[x]])
  }) %>%
  setNames(names(cms))
```

Calculations

```{r}
cell.df.rydbirk <- cell.df

# Cell-wise
cell.counts <- sapply(cms, rowSums) %>% 
  unname() %>% 
  unlist()

rem <- cell.counts %>%
  names() %>%
  table() %>%
  .[. > 1] %>%
  names()

cell.counts %<>% .[!names(.) %in% rem]

cell.group <- grepl.replace(names(cell.counts), patterns = c("CTRL","PD")) %>% as.factor()

mhc1.raw <- sapply(cms, function(x) {
  tmp <- x[,colnames(x) %in% c("A","B","C")]
  if(class(tmp) != "numeric") rowSums(tmp) else tmp
}) %>% 
  unname() %>% 
  unlist() %>%
  .[!names(.) %in% rem]

mhc2.raw <- sapply(cms, function(x) {
  tmp <- x[,!colnames(x) %in% c("A","B","C")]
  if(class(tmp) != "numeric") rowSums(tmp) else tmp
}) %>% 
  unname() %>% 
  unlist() %>%
  .[!names(.) %in% rem]

cell.df <- data.frame(group = cell.group,
                      counts = cell.counts,
                      mhc1 = mhc1.raw,
                      mhc2 = mhc2.raw) %>%
  mutate(., anno = anno[match(rownames(.), names(anno))],
         sample = rownames(.) %>% sapply(strsplit, "_", T) %>% sapply(`[[`, 1)) %>%
  `rownames<-`(names(cell.counts)) %>%
  filter(!is.na(anno))
```

Integrate with our data and plot

```{r, fig.width = 8, fig.height = 4}
cell.df.int <- rbind(cell.df.rydbirk %>% 
                       select(-c("depth","type")) %>%
                       mutate(origin = "Rydbirk"),
                     cell.df %>% 
                       mutate(anno = anno %>% factor(labels = c("Astrocytes","Exc. neurons","Inh. neurons","Endothelial","Eppendymal","Exc. neurons","Inh. neurons","Inh. neurons","Microglia","Oligodendrocytes","OPCs","Pericytes")),
                              origin = "Smajic")) %>%
  filter(! anno %in% c("Blood_immune","Endothelial","Eppendymal","Mural","Pericytes","Pericytes/endothelial","Immune")) %>%
  filter(!is.na(anno)) %>% 
  mutate(anno = anno %>% factor(),
         origin = origin %>% factor(),
         anno = anno %>% unname() %>% factor(),
         sample = sample %>% unname())

cell.anno.df <-
  cell.df.int %>%
  group_by(anno, sample, group, origin) %>%
  summarise(mhc1 = mean(mhc1),
            mhc2 = mean(mhc2),
            counts = mean(counts),
            cells = length(sample)) %>%
  as.data.frame() %>% 
  select(-cells, -counts, -sample)

p.text1 <- data.frame(signif = c("n.s.", "n.s.", "n.s.", "n.s.", "n.s.", "0.040", "0.032", "n.s."), 
                      mhc1 = 4.5,
                      anno = levels(cell.anno.df$anno))

p.text2 <- data.frame(signif = c("0.00057"), 
                      mhc2 = 8.5,
                      anno = "             Microglia")

cowplot::plot_grid(plotlist = list(
  ggplot(cell.anno.df, 
         aes(anno, 
             mhc1)) +
    geom_boxplot(outlier.shape = NA, 
                 aes(fill = group)) +
    geom_point(position = position_jitterdodge(jitter.width = 0.2), 
               aes(fill = group, 
                   col = origin, 
                   shape = group),
               size = 2) +
    scale_color_manual(values = c("black",
                                  "grey50")) +
    labs(x = "", 
         y = "Mean counts per sample", 
         title = "MHC-I counts", 
         fill = "", 
         col = "", 
         shape = " ") +
    geom_text(data = p.text1, aes(label = signif)) +
    theme_bw() +
    scale_fill_manual(values = pal.major) +
    theme(legend.position = "none", 
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          line = element_blank()),
  ggplot(cell.anno.df %>% filter(anno == "Microglia") %>% mutate(anno = "             Microglia"), aes(anno, mhc2)) +
    geom_boxplot(outlier.shape = NA, aes(fill = group)) +
    geom_point(position = position_jitterdodge(jitter.width = 0.2), aes(fill = group, col = origin, shape = group), size = 2) +
    scale_color_manual(values = c("black","grey50")) +
    labs(x = "", y = "", title = "MHC-II counts", fill = "", col = "", shape = " ") +
    geom_text(data = p.text2, aes(label = signif)) +
    theme_bw() +
    scale_fill_manual(values = pal.major) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          line = element_blank())
), ncol = 2, rel_widths = c(2,0.9))
```

Kruskal-Wallis, asses differences per cell type

```{r}
cell.anno.df %>% 
  group_by(anno) %>% 
  kruskal_test(mhc1 ~ group) %>% 
  data.frame()

cell.anno.df %>% 
  filter(anno == "Microglia") %>% 
  group_by(anno) %>% 
  kruskal_test(mhc2 ~ group) %>% 
  data.frame()
```

Wilcoxon, assess differences between our data and Smajic per condition per cell type

```{r}
cell.anno.df %>% 
  filter(group != "MSA", 
         !anno %in% c("MSN","PVMs")) %>%
  group_by(anno, group) %>% 
  wilcox_test(mhc1 ~ origin) %>% 
  data.frame()

cell.anno.df %>% 
  filter(group != "MSA", 
         anno == "Microglia") %>%
  group_by(anno, group) %>% 
  wilcox_test(mhc2 ~ origin) %>% 
  data.frame()
```

## 9b

We provide a curated list with MHC and cytokine genes. Please note, the order of clusters may change.

```{r, fig.width=4, fig.height=4}
dat <- read.table("MHC_cytokines_curated.tsv", header = T, sep = "\t")

cm.merged <- con$getJointCountMatrix(raw = T) %>% 
  Matrix::t() %>% 
  .[, colnames(.) %in% names(anno.major)] %>% 
  .[rownames(.) %in% (dat$name[dat$class == "cytokine"] %>% gsub("-","",.)), ] %>% 
  .[rowSums(.) > 0,]

# Create sample-wise annotation
anno.donor <- con$getDatasetPerCell()[colnames(cm.merged)]
anno.subtype <- anno.major %>% 
  .[!is.na(.) & (. %in% c("Microglia", "PVMs"))] %>% 
  factor()

idx <- intersect(anno.donor %>% names(), anno.subtype %>% names())

anno.donor %<>% .[idx]
anno.subtype %<>% .[idx] %>% 
  {`names<-`(as.character(.), names(.))}

anno.final <- paste0(anno.donor,"!!",anno.subtype) %>% 
  `names<-`(anno.donor %>% names())

# Create pseudo CM
cm.pseudo <- sccore::collapseCellsByType(cm.merged %>% Matrix::t(), 
                                         groups = anno.final, min.cell.count = 1) %>% 
  t() %>%
  apply(2, as, "integer")

cm.pseudo %<>% 
  {. + 1} %>% 
  DESeq2::DESeqDataSetFromMatrix(., 
                                 colnames(.) %>% 
                                   strsplit("_") %>% 
                                   sget(1) %>% 
                                   data.frame() %>% 
                                   `dimnames<-`(list(colnames(cm.pseudo), "group")), 
                                 design = ~ group) %>% 
  DESeq2::estimateSizeFactors() %>% 
  DESeq2::counts(normalized = T)

cm.pseudo %<>% 
  Matrix::t() %>% 
  scale() %>%
  Matrix::t()

ord <- cm.pseudo %>% 
  colnames() %>% 
  {data.frame(id = .)} %>% 
  mutate(., 
         ct = strsplit(id, "!!") %>% 
           sget(2),
         condition = strsplit(id, "!!|_") %>% 
           sget(1)) %>% 
  arrange(ct, condition) %>% 
  pull(id)

cm.pseudo %<>% 
  .[, match(ord, colnames(.))]

cm.pseudo %<>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "gene") %>% 
  reshape2::melt(id.vars = c("gene")) %>% 
  mutate(variable = as.character(variable),
         condition = strsplit(variable, "!!|_") %>% 
           sget(1),
         ct = strsplit(variable, "!!") %>% 
           sget(2)) %>% 
  group_by(condition, ct, gene) %>% 
  summarize(m = mean(value)) %>% 
  ungroup() %>% 
  mutate(variable = paste(condition, ct, sep = "!!")) %>% 
  select(-condition, -ct) %>% 
  reshape2::dcast(gene ~ variable, value.var = "m") %>% 
  tibble::column_to_rownames(var = "gene")

tmp <- cm.pseudo %>% 
  colnames() %>% 
  strsplit("_|!!")

clusters <- tmp %>%
  sget(2)

condition <- tmp %>% 
  sget(1)

ha <- ComplexHeatmap::HeatmapAnnotation(Annotation = clusters,
                                        Condition = condition,
                                        col = list(Annotation = c("Microglia" = unname(pal.major["Microglia"]), 
                                                                  "PVMs" = unname(pal.major["PVMs"])),
                                                   Condition = c("CTRL" = unname(pal.major["CTRL"]), 
                                                                 "MSA" = unname(pal.major["MSA"]),
                                                                 "PD" = unname(pal.major["PD"]))
                                        ))

ComplexHeatmap::Heatmap(cm.pseudo,
                        name = "Expression", 
                        show_column_names = F, 
                        show_row_names = T, 
                        cluster_columns = F, 
                        show_column_dend = F, 
                        cluster_rows = T, 
                        top_annotation = ha, 
                        show_row_dend = F, 
                        column_split = colnames(cm.pseudo) %>% strsplit("!!|_") %>% sget(2) %>% unname() %>% factor(),
                        col=circlize::colorRamp2(c(-1, 1.2), c("white", "firebrick")),
                        row_km = 4)
```

# Extended Data Figure 10

## 10c

```{r, fig.width=14, fig.height=6}
c("FOSL2", "TCF4", "STAT1", "NR3C1", "ETV7", "THRB", "BPTF", "RELB", "CEBPD", "HIF1A", "ELF1", "CREM", "ELF2", "ZNF44", "CHD1", "POU2F1", "GTF2IRD1", "NFIA", "NFE2L2", "FOXP1", "FOXO3") %>% 
  clusterProfiler::enrichGO("org.Hs.eg.db", "SYMBOL", "BP") %>% 
  enrichplot::pairwise_termsim() %>% 
  enrichplot::treeplot() +
  scale_fill_manual(values = brewer.pal(6, "Greys")[-1])
```

# Extended Data Figure 11

## 11a

We provide the results from scDRS based on Nalls *et al.* GWAS summary stats.

```{r, fig.width=6, fig.height=4}
dat.scores <- qread("gwas/nalls/PD.full_score.qs")

# Load downstream analyses for cell types versus trait
dat.ct <- read.delim("gwas/nalls/downstream_celltype.tsv", header = F, sep = "\t") %$%  
  strsplit(V1, " ") %>% 
  unlist() %>% 
  gsub("\\", "", ., fixed = T) %>% 
  gsub("\n", "", ., fixed = T) %>% 
  .[!sapply(., \(x) x == "")] %>% 
  .[c(1:5,69:72, # Header
      8:12,75:78, # Astro
      15:19,81:84, # EN
      21:25,86:89, # Immune
      28:32,92:95, # IN
      34:38,97:100, # MSN
      40:44,102:105, # MIC
      46:50,107:110, # OPCs
      52:56,112:115, # OL
      58:62,117:120, # PVMs
      64:68,122:125 # Peri
  )]

dat.sign <- dat.ct %>%
  split(seq(9)) %>% 
  bind_rows() %>% 
  {`colnames<-`(.[-1, ], .[1, ])} %>% 
  as.data.frame()

dat.sign %<>%  mutate(group = c("Astrocytes",
                                "Exc. neurons",
                                "Immune",
                                "Inh. neurons",
                                "MSN",
                                "Microglia",
                                "OPCs",
                                "Oligodendrocytes",
                                "PVMs",
                                "Pericytes/endothelial"))

dat.sign %<>% 
  filter(!group == "Unknown") %>% 
  mutate(p.adj = p.adjust(assoc_mcp, method = "fdr"),
         hetero.adj = p.adjust(hetero_mcp, method = "fdr")) %>% 
  mutate(assoc_sign = sapply(assoc_mcp, \(x) if (x <= 0.001) "***" else if (x <= 0.01 && x > 0.001) "**" else if (x <= 0.05 && x > 0.01) "*" else ""),
         hetero_sign = sapply(hetero_mcp, \(x) if (x <= 0.001) "###" else if (x <= 0.01 && x > 0.001) "##" else if (x <= 0.05 && x > 0.01) "#" else " ")) %>% 
  mutate(sign = paste0(assoc_sign,",",hetero_sign) %>% gsub(", ", "", .)) %>% 
  select(group, sign) %>% 
  dplyr::rename(type = group)

data.frame(cell = dat.scores$X,
           score = dat.scores$norm_score,
           type = anno.major[match(dat.scores$X,
                                   names(anno.major))]) %>%
  group_by(type) %>%
  summarize(m = median(score)) %>%
  filter(!is.na(type)) %>% 
  mutate(type = as.character(type)) %$% 
  arrange(., desc(m)) %>%
  mutate(type = factor(type, levels = type), sign = dat.sign$sign[match(type %>% levels, dat.sign$type)]) %>%
  ggplot(aes(type, m, fill = type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sign), vjust = -0.5) +
  theme_bw() + 
  labs(y = "Mean score", x = "", title = "scDRS mean scores and associations") +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        line = element_blank()) +
  ylim(c(-0.6, 2)) +
  scale_fill_manual(values = pal.major) +
  geom_hline(yintercept = 0)
```

## 11b

```{r, fig.height=4, fig.width=3.5}
dat.ct <- read.delim("gwas/nalls/downstream_celltype_microglia.tsv", header = F, sep = "\t") %$%  
  strsplit(V1, " ") %>% 
  unlist() %>% 
  gsub("\\", "", ., fixed = T) %>% 
  gsub("\n", "", ., fixed = T) %>% 
  gsub("group", "", .) %>% 
  .[!sapply(., \(x) x == "")] %>% 
  {c("group", .[seq(35)])} %>% 
  matrix(ncol = 6, byrow = T) %>% 
  as.data.frame() %>% 
  `colnames<-`(., .[1, ]) %>% 
  .[-1, ]

dat.sign <- 
  dat.ct %>% 
  filter(!group == "Unknown") %>% 
  mutate(p.adj = p.adjust(assoc_mcp, method = "fdr"),
         hetero.adj = p.adjust(hetero_mcp, method = "fdr")) %>% 
  mutate(assoc_sign = sapply(assoc_mcp, \(x) if (x <= 0.001) "***" else if (x <= 0.01 && x > 0.001) "**" else if (x <= 0.05 && x > 0.01) "*" else ""),
         hetero_sign = sapply(hetero_mcp, \(x) if (x <= 0.001) "###" else if (x <= 0.01 && x > 0.001) "##" else if (x <= 0.05 && x > 0.01) "#" else " ")) %>% 
  mutate(sign = paste0(assoc_sign,",",hetero_sign) %>% gsub(", ", "", .)) %>% 
  select(group, sign) %>% 
  dplyr::rename(type = group)

data.frame(cell = dat.scores$X %>% gsub("one_", "!!", .),
           score = dat.scores$norm_score,
           type = anno.major[match(dat.scores$X,
                                   names(anno.major))]) %>%
  filter(cell %in% names(anno.micro)) %>% 
  mutate(type = factor(anno.micro[.$cell])) %>% 
  group_by(type) %>%
  summarize(m = mean(score)) %>%
  filter(!is.na(type)) %>% 
  mutate(type = as.character(type)) %>% 
  arrange(desc(m)) %>% 
  mutate(type = factor(type, levels = type), sign = dat.sign$sign[match(type %>% levels, dat.sign$type)]) %>%
  mutate(type = factor(type, levels = type)) %>% 
  ggplot(aes(type, m, fill = type)) + 
  geom_bar(stat = "identity") +
  geom_text(aes(label = sign), vjust = -0.5, nudge_y = -0.05) +
  theme_bw() + 
  theme(line = element_blank()) +
  labs(y = "Mean score", x = "", title = "scDRS for microglia") +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ylim(c(0, 2)) +
  scale_fill_manual(values = pal.major)
```

## 11c

We provide the results from scDRS based on Chia *et al.* GWAS summary stats.

```{r, fig.width=6, fig.height=4}
dat.scores <- qread("gwas/chia/MSA.full_score.qs")

# Load downstream analyses for cell types versus trait
dat.ct <- read.delim("gwas/chia/downstream_celltype.tsv", header = F, sep = "\t") %$%  
  strsplit(V1, " ") %>% 
  unlist() %>% 
  gsub("\\", "", ., fixed = T) %>% 
  gsub("\n", "", ., fixed = T) %>% 
  .[!sapply(., \(x) x == "")] %>% 
  .[c(1:5,69:72, # Header
      8:12,75:78, # Astro
      15:19,81:84, # EN
      21:25,86:89, # Immune
      28:32,92:95, # IN
      34:38,97:100, # MSN
      40:44,102:105, # MIC
      46:50,107:110, # OPCs
      52:56,112:115, # OL
      58:62,117:120, # PVMs
      64:68,122:125 # Peri
  )]

dat.sign <- dat.ct %>%
  split(seq(9)) %>% 
  bind_rows() %>% 
  {`colnames<-`(.[-1, ], .[1, ])} %>% 
  as.data.frame()

dat.sign %<>%  mutate(group = c("Astrocytes",
                                "Exc. neurons",
                                "Immune",
                                "Inh. neurons",
                                "MSN",
                                "Microglia",
                                "OPCs",
                                "Oligodendrocytes",
                                "PVMs",
                                "Pericytes/endothelial"))

dat.sign %<>% 
  filter(!group == "Unknown") %>% 
  mutate(assoc_sign = sapply(assoc_mcp, \(x) if (x <= 0.001) "***" else if (x <= 0.01 && x > 0.001) "**" else if (x <= 0.05 && x > 0.01) "*" else ""),
         hetero_sign = sapply(hetero_mcp, \(x) if (x <= 0.001) "###" else if (x <= 0.01 && x > 0.001) "##" else if (x <= 0.05 && x > 0.01) "#" else " ")) %>% 
  mutate(sign = paste0(assoc_sign,",",hetero_sign) %>% gsub(",", "", .)) %>% 
  select(group, sign) %>% 
  dplyr::rename(type = group)

data.frame(cell = dat.scores$X,
           score = dat.scores$norm_score,
           type = anno.major[match(dat.scores$X,
                                   names(anno.major))]) %>%
  group_by(type) %>%
  summarize(m = median(score)) %>%
  filter(!is.na(type)) %>% 
  mutate(type = as.character(type)) %$% 
  arrange(., desc(m)) %>%
  mutate(type = factor(type, levels = type), sign = dat.sign$sign[match(type %>% levels, dat.sign$type)]) %>%
  ggplot(aes(type, m, fill = type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sign), vjust = 1.5) +
  theme_bw() + 
  labs(y = "Mean score", x = "", title = "scDRS mean scores and associations") +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        line = element_blank()) +
  ylim(c(-0.5, 0.5)) +
  scale_fill_manual(values = pal.major) +
  geom_hline(yintercept = 0)
```

# Extended Data Figure 12

All subfigures were made in GraphPad Prism. Data are available in `Phagocytosis_assay_triplicates_raw_data_HH.tsv` (Copenhagen experiment) or `HOUSTON.tsv` (Houston experiment).

# Session info

Time to knit

```{r}
Sys.time() - tt
```

```{r}
sessionInfo()
```