### 2024-12-04 Objects_preparations.Rmd
Added `Objects_preparations.Rmd`

---
title: "Objects preparations"
author: "Rasmus Rydbirk"
date: "2024-11-29"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Setup

```{r setup, message=FALSE}
library(qs)
library(cacoa)
library(conos)
library(magrittr)

anno.major <- qread("anno_major.qs")
```

# All cells

## Conos object

The easiest way to extract the list with raw count matrices is to extract it from the Conos object that we provide for all cells. We do this first and then preprocess.

```{r}
con.major <- qread("con_major.qs", nthreads = 10)

cms.preprocessed <- con.major$samples %>% 
  lget("misc") %>% 
  lget("rawCounts") %>% 
  lapply(basicP2proc, make.geneknn = F, get.tsne = F, get.largevis = F, n.cores = 32)
```

Then, we can re-create the Conos object. Please note, we provide an additional embedding in the prepared object that you'll lose if you save this object. Also, the embedding may not be exactly the same.

After the next step, we use the major object to create objects for subpopulations.

```{r, eval = F}
con.major <- cms.preprocessed %>%  
  Conos$new(n.cores = 32)

con.major$buildGraph()
con.major$findCommunities(min.group.size = 25)
con.major$embedGraph(method = "UMAP")

qsave("con_major.qs", nthreads = 10)
```

## Export to Python

```{r}
annotation <- qread("anno_major.qs")

con.major$samples %>% 
  lapply(conos:::getGenes) %>% 
  Reduce(union, .) %>% 
  .[grep("MT-|RPS|RPL", .)] %>% 
  prepareObjectsForPython(con.major, 
                          annotation,
                          "",
                          "all_cells",
                          "UMAP", 
                          genes.to.omit = .)
```

Subtypes

```{r}
anno.sub <- c("anno_astro.qs",
              "anno_micro_pvm.qs",
              "anno_neurons.qs",
              "anno_oligo.qs",
              "anno_opc.qs",
              "anno_pericytes.qs") %>% 
  lapply(qread) %>% 
  Reduce(c, .) %>% 
  .[!is.na(.)] %>%
  factor() %>% 
  renameAnnotation("Reactive_astrocytes","AS_reactive") %>% 
  renameAnnotation("Homeostatic_astrocytes","AS_homeostatic") %>% 
  renameAnnotation("Steady-state","MIC_steady-state") %>% 
  renameAnnotation("Intermediate1","MIC_intermediate1") %>% 
  renameAnnotation("Intermediate2","MIC_intermediate2") %>% 
  renameAnnotation("Activated","MIC_activated") %>% 
  renameAnnotation("Homeostatic_LINC01608","OL_LINC01608") %>% 
  renameAnnotation("Homeostatic_SLC5A11","OL_SLC5A11") %>% 
  renameAnnotation("Reactive_SCGZ","OL_SGCZ")

anno.df <- anno.sub %>% {
  data.frame(cellid = names(.), annotation = unname(.))
}

write.table(anno.df, 
            "all_cells_subtypes.annotation",
            sep = ",",
            dec = ".", 
            row.names = F,
            col.names = F)
```

## Cacoa objects

Create objects

CTRL vs. MSA

```{r, eval = F}
con_msa <- Conos$new(con$samples[!grepl("PD",names(con$samples))])

sample.groups <- con_msa$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "MSA"), .)

con_msa$buildGraph()
con_msa$findCommunities()
con_msa$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_msa, sample.groups, anno.major, ref.level = "CTRL", target.level = "MSA", n.cores = 50)

cao$estimateCellLoadings()
cao$estimateExpressionShiftMagnitudes(cell.groups = anno.major[!anno.major %in% c("Immune","Pericytes/endothelial")] %>% factor(), min.cells.per.sample = 3)

qsave(cao, "cao_major_msa.qs", nthreads = 10)
```

CTRL vs. PD

```{r, eval = F}
con_pd <- Conos$new(con$samples[!grepl("MSA",names(con$samples))])

sample.groups <- con_pd$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "PD"), .)

con_pd$buildGraph()
con_pd$findCommunities()
con_pd$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_pd, sample.groups, anno.major, ref.level = "CTRL", target.level = "PD", n.cores = 50)

cao$estimateCellLoadings()
cao$estimateExpressionShiftMagnitudes(cell.groups = anno.major[!anno.major %in% c("Immune","Pericytes/endothelial")] %>% factor(), min.cells.per.sample = 3)

qsave(cao, "cao_major_pd.qs", nthreads = 10)
```

PD vs. MSA

```{r, eval = F}
con_dis <- Conos$new(con$samples[!grepl("CTRL",names(con$samples))])

sample.groups <- con_dis$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("MSA", .), "MSA", "PD"), .)

con_dis$buildGraph()
con_dis$findCommunities()
con_dis$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_dis, sample.groups, anno.major, ref.level = "PD", target.level = "MSA", n.cores = 50)

cao$estimateCellLoadings()
cao_dis$estimateExpressionShiftMagnitudes(cell.groups = anno.major[!anno.major %in% c("Immune","Pericytes/endothelial")] %>% factor(), min.cells.per.sample = 3)

qsave(cao, "cao_major_dis.qs", nthreads = 10)
```

## Create count matrices per cell type

At this point, we can create lists of count matrices per cell type. It's not strictly necessary, but it was the approach we used.

```{r}
cell.ids <- anno.major %>%
  levels() %>% 
  lapply(\(ct) anno.major[anno.major == ct] %>% 
           names()) %>% 
  setNames(anno.major %>% levels())
```

```{r}
cms.all <- con.major$samples %>% 
  lget("misc") %>% 
  lget("rawCounts")

cms.per.ct <- cell.ids %>% 
  plapply(\(ct) {
    cms.all %>% 
      lapply(\(cm) cm[,colnames(cm) %in% ct])
  }, n.cores = 8) %>% 
  setNames(cell.ids %>% names())
```

# Neurons

## Conos

Remove samples with less than 25 cells

```{r}
cms <- cms.per.ct[["Neurons"]] %>% 
  .[sapply(., ncol) > 25]
```

```{r}
cms.preprocessed <- lapply(cms, 
                           basicP2proc, 
                           n.cores = 50, 
                           min.cells.per.gene = 10, 
                           get.largevis = FALSE, 
                           get.tsne = FALSE, 
                           make.geneknn = FALSE,
                           nPcs = 10, 
                           k = 20)

con <- Conos$new(cms.preprocessed, 
                 n.cores = 32, 
                 verbose = TRUE)

con$buildGraph(ncomps = 20)
con$findCommunities()
con$embedGraph(method = "UMAP")
```

## Cacoa

CTRL vs combined disease

```{r}
sample.groups <- con$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "DISEASE"), .)

anno <- qread("anno_neurons.qs")

cao <- Cacoa$new(con, sample.groups, anno, ref.level = "CTRL", target.level = "DISEASE", n.cores = 50)

qsave(cao, "cao_neurons.qs", nthreads = 10)
```

CTRL vs MSA

```{r}
con_msa <- Conos$new(con$samples[!grepl("PD",names(con$samples))])

sample.groups <- con_msa$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "MSA"), .)

con_msa$buildGraph()
con_msa$findCommunities()
con_msa$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_msa, sample.groups, anno, ref.level = "CTRL", target.level = "MSA", n.cores = 50)

cao$estimateCellLoadings()
cao$estimateExpressionShiftMagnitudes(min.cells.per.sample = 1)

qsave(cao, "cao_neurons_msa.qs", nthreads = 10)
```

CTRL vs PD

```{r, eval = F}
con_pd <- Conos$new(con$samples[!grepl("MSA",names(con$samples))])

sample.groups <- con_pd$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "PD"), .)

con_pd$buildGraph()
con_pd$findCommunities()
con_pd$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_pd, sample.groups, anno, ref.level = "CTRL", target.level = "PD", n.cores = 50)

cao$estimateCellLoadings()
cao$estimateExpressionShiftMagnitudes(min.cells.per.sample = 1)

qsave(cao, "cao_neurons_pd.qs", nthreads = 10)
```

PD vs MSA

```{r, eval = F}
con_dis <- Conos$new(con$samples[!grepl("CTRL",names(con$samples))])

sample.groups <- con_dis$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("MSA", .), "MSA", "PD"), .)

con_dis$buildGraph()
con_dis$findCommunities()
con_dis$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_dis, sample.groups, anno, ref.level = "PD", target.level = "MSA", n.cores = 50)

cao$estimateCellLoadings()
cao$estimateExpressionShiftMagnitudes(min.cells.per.sample = 1)

qsave(cao, "cao_neurons_dis.qs", nthreads = 10)
```

# Glia

## Conos

```{r}
cell.ids.glia <- anno.major[anno.major %in% c("Oligodendrocytes","Astrocytes","OPCs")] %>% 
  names()

cms.glia <- cms.all %>% 
  lapply(\(cm) cm[,colnames(cm) %in% cell.ids.glia]) %>% 
  setNames(cms.all %>% names())

cms.preprocessed <- lapply(cms.glia, 
                           basicP2proc, 
                           n.cores = 50, 
                           min.cells.per.gene = 10, 
                           get.largevis = FALSE, 
                           get.tsne = FALSE, 
                           make.geneknn = FALSE)

con <- Conos$new(cms.preprocessed, 
                 n.cores = 100, 
                 verbose = TRUE)

con$buildGraph(space = "CPCA", alignment.strength = 0.1)
con$findCommunities()
con$embedGraph(method = "largeVis", embedding.name = "largeVis_CPCA_AS01")

qsave(con, "con_oligo_astro_opc.qs", nthreads = 10)
```

## Cacoa

CTRL vs MSA

```{r, eval = F}
con_msa <- Conos$new(con$samples[!grepl("PD",names(con$samples))])

sample.groups <- con_msa$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "MSA"), .)

con_msa$buildGraph()
con_msa$findCommunities()
con_msa$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_msa, sample.groups, anno, ref.level = "CTRL", target.level = "MSA", n.cores = 32)

cao$estimateCellLoadings()
cao$estimateExpressionShiftMagnitudes()

qsave(cao, "cao_oligo_astro_opc_msa.qs", nthreads = 10)
```

CTRL vs PD

```{r, eval = F}
con_pd <- Conos$new(con$samples[!grepl("MSA",names(con$samples))])

sample.groups <- con_pd$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "PD"), .)

con_pd$buildGraph()
con_pd$findCommunities()
con_pd$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_pd, sample.groups, anno, ref.level = "CTRL", target.level = "PD", n.cores = 32)

cao$estimateCellLoadings()
cao$estimateExpressionShiftMagnitudes()

qsave(cao, "cao_oligo_astro_opc_pd.qs", nthreads = 10)
```

PD vs MSA

```{r, eval = F}
con_dis <- Conos$new(con$samples[!grepl("CTRL",names(con$samples))])

sample.groups <- con_dis$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("PD", .), "PD", "MSA"), .)

con_dis$buildGraph()
con_dis$findCommunities()
con_dis$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_dis, sample.groups, anno, ref.level = "PD", target.level = "MSA", n.cores = 32)

cao$estimateCellLoadings()
cao$estimateExpressionShiftMagnitudes()

qsave(cao, "cao_oligo_astro_opc_dis.qs", nthreads = 10)
```

# Oligodendrocytes

## Conos

```{r}
cms <- cms.per.ct[["Oligodendrocytes"]]

cms.preprocessed <- lapply(cms, 
                           basicP2proc, 
                           n.cores = 50, 
                           min.cells.per.gene = 10, 
                           get.largevis = FALSE, 
                           get.tsne = FALSE, 
                           make.geneknn = FALSE)

con <- Conos$new(cms.preprocessed, 
                 n.cores = 100, 
                 verbose = TRUE)

con$buildGraph(space = "CPCA")
con$findCommunities()
con$embedGraph(method = "largeVis", embedding.name = "largeVis_CPCA")
```

## Cacoa

CTRL vs MSA

```{r}
con_msa <- Conos$new(con$samples[!grepl("PD",names(con$samples))])

sample.groups <- con_msa$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "MSA"), .)

con_msa$buildGraph()
con_msa$findCommunities()
con_msa$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_msa, sample.groups, anno, ref.level = "CTRL", target.level = "MSA", n.cores = 50)

## At this stage we'll remove mitochondrial and ribosomal genes since they skew DE/GO results, and they are ambient/unreal in single nucleus data.

all.genes <- cao$data.object$samples %>% 
  lget("misc") %>% 
  lget("rawCounts") %>% 
  lapply(colnames) %>% 
  Reduce(union, .)

genes.to.omit <- all.genes %>% 
  .[grepl("MT-|RPL|RPS", .)]

cao$estimateDEPerCellType(resampling.method = "bootstrap", genes.to.omit = genes.to.omit)
cao$estimateOntology(type = "GSEA", 
                     org.db = org.Hs.eg.db::org.Hs.eg.db)
cao$estimateOntologyFamilies(name = "GSEA")

qsave(cao, "cao_oligo_msa.qs", nthreads = 10)
```

CTRL vs PD

```{r}
con_pd <- Conos$new(con$samples[!grepl("MSA",names(con$samples))])

sample.groups <- con_pd$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "PD"), .)

con_pd$buildGraph()
con_pd$findCommunities()
con_pd$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_pd, sample.groups, anno, ref.level = "CTRL", target.level = "PD", n.cores = 50)

## At this stage we'll remove mitochondrial and ribosomal genes since they skew DE/GO results, and they are ambient/unreal in single nucleus data.

all.genes <- cao$data.object$samples %>% 
  lget("misc") %>% 
  lget("rawCounts") %>% 
  lapply(colnames) %>% 
  Reduce(union, .)

genes.to.omit <- all.genes %>% 
  .[grepl("MT-|RPL|RPS", .)]

cao$estimateDEPerCellType(resampling.method = "bootstrap", genes.to.omit = genes.to.omit)
cao$estimateOntology(type = "GSEA", 
                     org.db = org.Hs.eg.db::org.Hs.eg.db)
cao$estimateOntologyFamilies(name = "GSEA")

qsave(cao, "cao_oligo_pd.qs", nthreads = 10)
```

PD vs MSA

```{r}
con_dis <- Conos$new(con$samples[!grepl("CTRL",names(con$samples))])

sample.groups <- con_dis$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("MSA", .), "MSA", "PD"), .)

con_dis$buildGraph()
con_dis$findCommunities()
con_dis$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_dis, sample.groups, anno, ref.level = "PD", target.level = "MSA", n.cores = 50)

## At this stage we'll remove mitochondrial and ribosomal genes since they skew DE/GO results, and they are ambient/unreal in single nucleus data.

all.genes <- cao$data.object$samples %>% 
  lget("misc") %>% 
  lget("rawCounts") %>% 
  lapply(colnames) %>% 
  Reduce(union, .)

genes.to.omit <- all.genes %>% 
  .[grepl("MT-|RPL|RPS", .)]

cao$estimateDEPerCellType(resampling.method = "bootstrap", genes.to.omit = genes.to.omit)
cao$estimateOntology(type = "GSEA", 
                     org.db = org.Hs.eg.db::org.Hs.eg.db)
cao$estimateOntologyFamilies(name = "GSEA")

qsave(cao, "cao_oligo_dis.qs", nthreads = 10)
```

# Astrocytes

## Conos

```{r}
cms <- cms.per.ct[["Astrocytes"]]

cms.preprocessed <- lapply(cms, 
                           basicP2proc, 
                           n.cores = 50, 
                           min.cells.per.gene = 10, 
                           get.largevis = FALSE, 
                           get.tsne = FALSE, 
                           make.geneknn = FALSE,
                           nPcs = 50)

con <- Conos$new(cms.preprocessed, 
                 n.cores = 100, 
                 verbose = TRUE)

con$buildGraph(space = "CPCA")
con$findCommunities()
con$embedGraph(method = "largeVis")
```

## Export to Python

```{r}
annotation <- qread("anno_astro.qs")

con$samples %>% 
  lapply(conos:::getGenes) %>% 
  Reduce(union, .) %>% 
  .[grep("MT-|RPS|RPL", .)] %>% 
  prepareObjectsForPython(con, 
                          annotation,
                          "",
                          "astrocytes",
                          "largeVis_CPCA",
                          genes.to.omit = .)
```

## Cacoa

CTRL vs MSA

```{r}
con_msa <- Conos$new(con$samples[!grepl("PD",names(con$samples))])

sample.groups <- con_msa$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "MSA"), .)

con_msa$buildGraph()
con_msa$findCommunities()
con_msa$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_msa, sample.groups, anno, ref.level = "CTRL", target.level = "MSA", n.cores = 50)

## At this stage we'll remove mitochondrial and ribosomal genes since they skew DE/GO results, and they are ambient/unreal in single nucleus data.

all.genes <- cao$data.object$samples %>% 
  lget("misc") %>% 
  lget("rawCounts") %>% 
  lapply(colnames) %>% 
  Reduce(union, .)

genes.to.omit <- all.genes %>% 
  .[grepl("MT-|RPL|RPS", .)]

cao$estimateDEPerCellType(resampling.method = "bootstrap", genes.to.omit = genes.to.omit)
cao$estimateOntology(type = "GSEA", 
                     org.db = org.Hs.eg.db::org.Hs.eg.db)
cao$estimateOntologyFamilies(name = "GSEA")

qsave(cao, "cao_astro_msa.qs", nthreads = 10)
```

CTRL vs PD

```{r, eval = F}
con_pd <- Conos$new(con$samples[!grepl("MSA",names(con$samples))])

sample.groups <- con_pd$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "PD"), .)

con_pd$buildGraph()
con_pd$findCommunities()
con_pd$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_pd, sample.groups, anno, ref.level = "CTRL", target.level = "PD", n.cores = 50)

## At this stage we'll remove mitochondrial and ribosomal genes since they skew DE/GO results, and they are ambient/unreal in single nucleus data.

all.genes <- cao$data.object$samples %>% 
  lget("misc") %>% 
  lget("rawCounts") %>% 
  lapply(colnames) %>% 
  Reduce(union, .)

genes.to.omit <- all.genes %>% 
  .[grepl("MT-|RPL|RPS", .)]

cao$estimateDEPerCellType(resampling.method = "bootstrap", genes.to.omit = genes.to.omit)
cao$estimateOntology(type = "GSEA", 
                     org.db = org.Hs.eg.db::org.Hs.eg.db)
cao$estimateOntologyFamilies(name = "GSEA")

qsave(cao, "cao_astro_pd.qs", nthreads = 10)
```

PD vs MSA

```{r, eval = F}
con_dis <- Conos$new(con$samples[!grepl("CTRL",names(con$samples))])

sample.groups <- con_dis$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("MSA", .), "MSA", "PD"), .)

con_dis$buildGraph()
con_dis$findCommunities()
con_dis$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_dis, sample.groups, anno, ref.level = "PD", target.level = "MSA", n.cores = 50)

## At this stage we'll remove mitochondrial and ribosomal genes since they skew DE/GO results, and they are ambient/unreal in single nucleus data.

all.genes <- cao$data.object$samples %>% 
  lget("misc") %>% 
  lget("rawCounts") %>% 
  lapply(colnames) %>% 
  Reduce(union, .)

genes.to.omit <- all.genes %>% 
  .[grepl("MT-|RPL|RPS", .)]

cao$estimateDEPerCellType(resampling.method = "bootstrap", genes.to.omit = genes.to.omit)
cao$estimateOntology(type = "GSEA", 
                     org.db = org.Hs.eg.db::org.Hs.eg.db)
cao$estimateOntologyFamilies(name = "GSEA")

qsave(cao, "cao_astro_dis.qs", nthreads = 10)
```

# Microglia and PVMs

## Conos

```{r}
cell.ids.micro_pmv <- anno.major[anno.major %in% c("PVMs","Microglia")] %>% 
  names()

cms.micro_pmv <- cms.all %>% 
  lapply(\(cm) cm[,colnames(cm) %in% cell.ids.micro_pmv]) %>% 
  setNames(cms.all %>% names())

cms.preprocessed <- lapply(cms.micro_pmv, 
                           basicP2proc, 
                           n.cores = 50, 
                           min.cells.per.gene = 10, 
                           get.largevis = FALSE, 
                           get.tsne = FALSE, 
                           make.geneknn = FALSE)

con <- Conos$new(cms.preprocessed, 
                 n.cores = 100, 
                 verbose = TRUE)

con$buildGraph(space = "CPCA")
con$findCommunities()
con$embedGraph(method = "largeVis", embedding.name = "largeVis_CPCA")
```

## Export for Python

```{r}
annotation <- qread("anno_micro_pvm.qs")

con$samples %>% 
  lapply(conos:::getGenes) %>% 
  Reduce(union, .) %>% 
  .[grep("MT-|RPS|RPL", .)] %>% 
  prepareObjectsForPython(con, 
                          annotation,
                          "",
                          "microglia_pvms",
                          "largeVis_CPCA", 
                          genes.to.omit = .)
```

## Cacoa

CTRL vs combined disease

```{r}
cao <- Cacoa$new(con, sample.groups, anno, ref.level = "CTRL", target.level = "DISEASE", n.cores = 50)

qsave(cao, "cao_micro_pvm.qs", nthreads = 10)
```

CTRL vs MSA

```{r}
con_msa <- Conos$new(con$samples[!grepl("PD",names(con$samples))])

sample.groups <- con_msa$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "MSA"), .)

con_msa$buildGraph()
con_msa$findCommunities()
con_msa$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_msa, sample.groups, anno, ref.level = "CTRL", target.level = "MSA", n.cores = 50)

cao$estimateCellLoadings()

## At this stage we'll remove mitochondrial and ribosomal genes since they skew DE/GO results, and they are ambient/unreal in single nucleus data.

all.genes <- cao$data.object$samples %>% 
  lget("misc") %>% 
  lget("rawCounts") %>% 
  lapply(colnames) %>% 
  Reduce(union, .)

genes.to.omit <- all.genes %>% 
  .[grepl("MT-|RPL|RPS", .)]

cao$estimateDEPerCellType(resampling.method = "bootstrap", genes.to.omit = genes.to.omit)
cao$estimateOntology(type = "GSEA", 
                     org.db = org.Hs.eg.db::org.Hs.eg.db)
cao$estimateOntologyFamilies(name = "GSEA")

qsave(cao, "cao_micro_pvm_msa.qs", nthreads = 10)
```

CTRL vs PD

```{r}
con_pd <- Conos$new(con$samples[!grepl("MSA",names(con$samples))])

sample.groups <- con_pd$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("CTRL", .), "CTRL", "PD"), .)

con_pd$buildGraph()
con_pd$findCommunities()
con_pd$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_pd, sample.groups, anno, ref.level = "CTRL", target.level = "PD", n.cores = 50)

cao$estimateCellLoadings()

## At this stage we'll remove mitochondrial and ribosomal genes since they skew DE/GO results, and they are ambient/unreal in single nucleus data.

all.genes <- cao$data.object$samples %>% 
  lget("misc") %>% 
  lget("rawCounts") %>% 
  lapply(colnames) %>% 
  Reduce(union, .)

genes.to.omit <- all.genes %>% 
  .[grepl("MT-|RPL|RPS", .)]

cao$estimateDEPerCellType(resampling.method = "bootstrap", genes.to.omit = genes.to.omit)
cao$estimateOntology(type = "GSEA", 
                     org.db = org.Hs.eg.db::org.Hs.eg.db)
cao$estimateOntologyFamilies(name = "GSEA")

qsave(cao, "cao_micro_pvm_pd.qs", nthreads = 10)
```

PD vs MSA

```{r}
con_dis <- Conos$new(con$samples[!grepl("CTRL",names(con$samples))])

sample.groups <- con_dis$samples %>% 
  names() %>% 
  setNames(ifelse(grepl("MSA", .), "MSA", "PD"), .)

con_dis$buildGraph()
con_dis$findCommunities()
con_dis$embedGraph(method = "largeVis")

cao <- Cacoa$new(con_dis, sample.groups, anno, ref.level = "PD", target.level = "MSA", n.cores = 50)

cao$estimateCellLoadings()

## At this stage we'll remove mitochondrial and ribosomal genes since they skew DE/GO results, and they are ambient/unreal in single nucleus data.

all.genes <- cao$data.object$samples %>% 
  lget("misc") %>% 
  lget("rawCounts") %>% 
  lapply(colnames) %>% 
  Reduce(union, .)

genes.to.omit <- all.genes %>% 
  .[grepl("MT-|RPL|RPS", .)]

cao$estimateDEPerCellType(resampling.method = "bootstrap", genes.to.omit = genes.to.omit)
cao$estimateOntology(type = "GSEA", 
                     org.db = org.Hs.eg.db::org.Hs.eg.db)
cao$estimateOntologyFamilies(name = "GSEA")

qsave(cao, "cao_micro_pvm_dis.qs", nthreads = 10)
```